### 2017-01-11 paper.tex
initial commit of new paper draft

\documentclass[a4paper,titlepage]{scrartcl}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage[format=hang,margin=1cm]{caption}
\usepackage{rotating}
\usepackage{subcaption}
\usepackage{authblk}
\usepackage{color}
\usepackage{xspace}
\usepackage{textgreek}

\begin{document}

\input{definitions}

\sisetup{range-phrase=--}
\sisetup{range-units=single}

% ----------------------------------------------------------------------------
% Titlepage Parameters
% ----------------------------------------------------------------------------

\title{Arterial pulsations are not powerful enough to drive perivascular drainage of solutes from the brain}
\author[1,2]{Alexandra K. Diem}
\author[2]{Neil W. Bressloff}
\author[3]{Roxana O. Carare}
\author[4]{Giles Richardson}
\affil[1]{Institute for Complex Systems Simulation, University of Southampton, Southampton SO17 1BJ, UK}
\affil[2]{Computational Engineering and Design, Faculty of Engineering \& the Environment, University of Southampton, Southampton SO17 1BJ, UK}
\affil[3]{Faculty of Medicine, University of Southampton, Southampton, UK}
\affil[4]{Faculty of Mathematical Sciences, University of Southampton, Southampton, UK}

\maketitle

\begin{abstract}
\textbf{Summary.} Alzheimer's Disease (AD) is the most common form of dementia and to date there is no preventive medication. Its development is closely related to the accumulation of the neuronal metabolite \Ab, which raises the question of how metabolites are eliminated from the brain. Unlike the rest of the body the brain does not have a lymphatic system and therefore has to eliminate metabolites differently. A connection between the pathology of AD and cerebrovascular diseases has been suggested and confirmed in various mice studies, which has lead to the assumption that interstitial fluid in the basement membranes of cerebral arteries provide the pathways for the perivascular drainage of \Ab. However, perivascular drainage occurs in the reverse direction to arterial blood flow, an effect that lacks an effective explanation. In this study we tested the well-accepted hypothesis that arterial pulsations drive perivascular drainage and additionally that a valve mechanism exists which ensure net reverse drainage. This was examined in terms of a  mathematical model of the  drainage mechanism. It is shown that a valve mechanism (such as directional permeability of the perivascular space) is necessary to achieve a net reverse flow. However, it is also shown that arterial pulsations are not strong enough to produce drainage velocities comparable to experimental observations. % 209 words

Word count: 209

\vspace{1cm}

\textbf{Keywords:} Alzheimer's disease, brain lymphatics, cerebrovascular flow
\end{abstract}

% ----------------------------------------------------------------------------
% Paper Body
% ----------------------------------------------------------------------------

% Introduction

Dementia, a term often used synonymously with Alzheimer's disease (AD), is a symptom of a variety of diseases affecting the human brain and does not constitute a part of normal ageing. Although it has been studied for over 100 years \cite{Selkoe2001}, to date there is no cure as the processes relevant to its onset and progression have not yet been fully understood. The pathology of AD comprises different characteristics, some of which are degeneration and death of neurons, synapse loss, neuroinflammation and the accumulation of waste product deposits \cite{Koffie2011}. The protein amyloid-$\beta$ (\Ab) is produced as part of the normal metabolism of healthy brains, but it plays a major role in the development of AD \cite{Haass1992}. Its accumulation in the walls of cerebral blood vessels as cerebral amyloid angiopathy (CAA) is closely associated with the onset and progression of AD \cite{Weller2006, Weller2009}. CAA strongly correlates with the progression of AD and the degree of dementia as it appears in \SIrange{90}{96}{\percent} of AD patients \cite{Weller2009a}.

In order to process metabolites, the brain requires a lymphatic system that transports waste products to the lymph nodes, where they are decomposed. While the rest of the body possesses a dedicated network of lymphatic vessels these do not exist in the brain \cite{Weller2010}. There is ample evidence that the basement membranes (BM) in the walls of cerebral arteries provide the clearance pathway for interstitial fluid (ISF) towards the surface of the brain, also referred to as perivascular drainage \cite{Weller2010,Carare2008,Hawkes2011,Morris2014}. Because it provides the pathways for perivascular drainage, the BM is also referred to as perivascular space (PVS) in this context. It is important not to confuse this PVS with the Virchow-Robin spaces, which are also often referred to as perivascular (or ``paravascular'') spaces. However, Virchow-Robin spaces are virtual spaces and have no role in \Ab drainage. Additionally, the BM does not actually constitute a space, but rather a fluid-filled protein matrix. Our understanding of the mechanism that drives perivascular drainage through the BM is still very limited. The evidence indicates that ISF enters the BM at the capillary level and flows towards the lymph nodes in the neck via cerebral arteries, but not veins. This suggests that ISF flow is driven in the reverse direction of the blood flow, a phenomenon that does not seem very plausible at first sight \cite{Carare2008}.

A widely accepted hypothesis for the driving mechanism of this reverse perivascular drainage of solutes from the brain is that it is driven by arterial pulsations \cite{Weller2009,Carare2008,Hawkes2011,Morris2014,Schley2006,Attems2011,Wang2011,Iliff2012,Asgari2015,Sharp2015,Weller2015a}. In this paper we tested this hypothesis by using a multi-scale modelling approach combining analytical and numerical methods. The BM is modelled as a porous medium using Darcy's law and a lubrication approximation. Flow inside the BM is driven by both the ISF pressure gradient and the deformation of the arterial wall, which are obtained from a numerical model of the middle cerebral artery (MCA). To guarantee reverse drainage (\ie in the opposite direction of the blood flow) a valve mechanism is implemented as a pressure gradient dependent permeablility. It is shown that this valve mechanism is neccesary to achieve net reverse drainage. The results show that this valve mechanism successfully ensures net reverse drainage inside the BM and is neccesary. However, due to the long wavelength of arterial pulsations the pressure gradients inside the BM were found to be too small to drive fluid flow at the expected velocities. The resulting velocity of the model was approximately two orders of magnitude smaller than values obtained experimentally \cite{Carare2008}. Therefore arterial pulsations are concluded to be insufficient to drive perivascular drainage along the BM of cerebral arteries.

\section{Porous medium model of lymphatic drainage in the BM}
\label{sec:diodemodel}

We propose a model of lymphatic drainage from the brain via the cerebral BM based on Darcy's law for flow in porous media. By exploiting the discrepancy of scales between the width of the BM and its length we approximate the flow by a one-dimensional lubrication model (see for example \cite{Ockendon1995}). Furthermore, we allow the permeability of the BM to depend on the pressure gradient of the ISF. Such a pressure gradient dependent permeability could, for instance, model proteins in the BM protein network that have structural properties that allow them to bend easily only in one direction, thereby acting as a valve, and thus provide a higher resistance to flow in the direction of the blood flow. A similar idea has been investigated in \cite{Sharp2015}, where proteins were explicitly modelled as cillia. However, here we model the active component of the BM generically in this model and thus allow for alternative explanations of the phenomenon.

The BM consists of a complex matrix of proteins (extracellular matrix, ECM) that can be interpreted as a porous medium through which soluble metabolites drain via the ISF. We model flow through this medium, in the standard fashion, by Darcy's law in a 2D cylindrical coordinate system. Darcy's law states that the flux $\q$ is related to the fluid pressure gradient $\boldsymbol{\nabla} p$ by
\beq\label{eq:darcy}
  \q = -K(p_z) \boldsymbol{\nabla} p
\eeq
where $K(p_z)$\footnote{Note that the permeability $K$ used here is the ratio of the intrinsic permeability $k$ of the ECM to the viscosity $\mu$ of ISF, that is $K = k / \mu$} is the pressure gradient dependent permeability of the ECM. We allow the permeability $K$ to be a function of the axial component of the ISF pressure gradient $\partial p(z,t) / \partial z = p_z(z,t)$. In order to obtain a net flux up the vessel, counter to the blood flow, we require this function to be increasing in $p_z$, representing the notion that it is harder to push a flow in the positive $z$-direction ($p_z<0$ and $K(p_z)$ small) than in the negative $z$-direction ($p_z>0$ and $K(p_z)$ large)
\begin{equation}
  K(p_z) = \begin{cases}
    K_0 & p_z > 0\\
    K_1 & p_z \leq 0.
  \end{cases}
\end{equation}

Because of the boundary conditions used we will continue working in terms of the fluid velocity $\vel$, which is related to $\q$ by the linear relationship $\vel = \gamma \, \q$, where $\gamma$ is the porosity of the medium. The governing equations in cylindrical coordinates are thus
\begin{gather}
  u \vect{e_z} + v \vect{e_r} = - \gamma K(p_z) \left( \vect{e_z} p_z + \vect{e_r} p_r \right) \label{eq:darcy_velocity}\\
  u_z + \frac{1}{r} \frac{\dd}{\dd r} \left( r v \right) = 0 \label{eq:continuity},
\end{gather}
where $\vect{e_z}$ and $\vect{e_r}$ describe the unit vectors in the $z$ and $r$ direction. The continuity of velocity is given in \eqref{eq:continuity}.

\begin{figure}
\centerline{\includegraphics{figures/pvs_model.pdf}}
\caption{Geometry of the perivascular drainage pathways model. The length $L$ of a typical artery in the brain can be assumed to be much greater than the width $W$ of the perivascular drainage pathways. Therefore we can approximate the flow as a lubrication model. The pulsation of the capillary is described by the function $h(z,t)$.\label{fig:pvs_model}}
\end{figure}

To determine the boundary conditions of the system consider the geometry in Figure \ref{fig:pvs_model}. Under the assumption that the vessel deformations are radially symmetric the positions of the inner and outer sheaths bounding the BM can be described by the two expressions $r = R_i(z,t)$ and $r = R_o(z,t)$, respectively. Assuming that the flows through the surfaces of the BM are negligible the following kinematic boundary condition applies
\beq \label{eq:kinematic}  
\frac{D}{D t} \left( r - R_i(z,t) \right) = 0 \quad \text{on } r = R_i(z,t)
\eeq
and analogously for $R_o(z,t)$.

Because the BM thickness is much smaller than the vessel length it is possible, on using the lubrication approximation, to derive a 1D model (in $z$) for the BM thickness $h(z,t)$; this takes the form
\beq
\frac{\dd}{\dd t} \left( R_i(z,t) \cdot h(z,t) \right) = \frac{\dd}{\dd z} \left( R_i(z,t) \cdot h(z,t) \cdot K(p_{z}) \cdot p_{z} \right),\label{eq:pvs_model}
\eeq
which describes the evolution of the BM thickness and can be solved using a finite difference approximation. Consider an infinitesimally small volume of BM $dV = 2\pi \cdot R_i(z,t) \cdot h(z,t) \cdot dz$. Then the difference between the flux into and out of the element $dV$ is equal to $\dd (dV)/ \dd t$, which can be written as
\begin{align}
2\pi \frac{\dd}{\dd t} \left( R_i(z,t) \cdot h(z,t) \right) &= \nonumber \\ 
2\pi \cdot R_i(z,t) \cdot h(z,t) & \cdot q(z,t) - 2\pi \cdot R_i(z + dz,t) \cdot h(z + dz,t) \cdot q(z + dz,t),
\end{align}
which is equal to the model equation \eqref{eq:pvs_model}. For a detailed derivation see Supplementary Information 1. The pressure $p(z,t)$ and the displacement of the BM wall $R_i(z,t)$ are, to a very good approximation, determined solely by the arterial pressure (\ie almost independent of the BM flow), see Supplementary Information 2 for more details. $R_i(z,t)$ and $p_z$ can be obtained from a numerical artery model described in the next section.

\section{Numerical Model of a Straight Arterial Section}

\begin{table}
\center
\caption{Parameters used to simulate the MCA including the first bifurcation.\label{tab:parameter}}
\begin{tabular}{ccc}
\toprule
Parameter & Value & Physical meaning\\
\midrule
\multicolumn{3}{l}{MCA}\\
\midrule
$R_u$ & \SI{0.14}{\centi\metre} & upstream radius\\
$R_d$ & \SI{0.14}{\centi\metre} & downstream radius\\
$L$ & \SI{7}{\centi\metre} & length\\
\midrule
\multicolumn{3}{l}{Left daughter vessel}\\
\midrule
$R_u$ & \SI{0.13}{\centi\metre} & upstream radius\\
$R_d$ & \SI{0.08}{\centi\metre} & downstream radius\\
$L$ & \SI{6.5}{\centi\metre} & length\\
\midrule
\multicolumn{3}{l}{Right daughter vessel}\\
\midrule
$R_u$ & \SI{0.13}{\centi\metre} & upstream radius\\
$R_d$ & \SI{0.08}{\centi\metre} & downstream radius\\
$L$ & \SI{6.4}{\centi\metre} & length\\
\midrule
\multicolumn{3}{l}{Simulation parameters}\\
\midrule
$k_1$ & \SI{2.0e7}{\gram\per\centi\metre\per\square\second} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
$k_2$ & \SI{-22.53}{\per\centi\metre} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
$k_3$ & \SI{8.65e5}{\gram\per\centi\metre\per\square\second} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
$R_1$ & \SI{14130}{\gram\per\centi\metre\tothe{4}\per\second} & First resistance element\\
$R_2$ & \SI{7200}{\gram\per\centi\metre\tothe{4}\per\second} & Second resistance element\\
$C_T$ & \SI{2.4752e-3}{\centi\metre\tothe{4}\square\second\per\gram} & Compliance element\\
$T$ & \SI{0.85}{\second} & cardiac cycle length\\
$t_c$ & 4 & number of cardiac cycles simulated\\
$\Delta t$ & \SI{1e-5}{\second} & time step size\\
$\Delta z$ & \SI{0.1}{\centi\metre} & spatial step size\\
\bottomrule
\end{tabular}
\end{table}

Following the geometry outlined in \autoref{sec:diodemodel} arteries are modelled as axisymmetric tubes in a cylindrical coordinate system with cross-sectional area $A(z,t) = \pi R(z,t)^2$ and flux $q(z,t)$. Following the derivations in \cite{Olufsen2000,Diem2016} arteries are represented by the conservation law
\begin{equation}
  \frac{\dd \boldsymbol{U}}{\dd t} + \frac{\dd \boldsymbol{F}}{\dd z} = \boldsymbol{S}, \label{eq:conservation}
\end{equation}
where
\begin{gather*}
  \boldsymbol{U} = \begin{pmatrix} A(z,t) \\ q(z,t) \end{pmatrix}, \quad \boldsymbol{F} = \begin{pmatrix} q(z,t)\\ \dfrac{q(z,t)^2}{A(z,t)} + f(r_0) \sqrt{A_0(z) A(z,t)} \end{pmatrix}, \quad \boldsymbol{S} = \begin{pmatrix} 0\\ S_1 \end{pmatrix}\\
  \begin{split}
    S_1 = -&\frac{2 \pi R(z,t)}{\delta_b \mathrm{Re}} \frac{q(z,t)}{A(z,t)} +\\
    &\left( 2 \sqrt{A(z,t)} \left( \sqrt{\pi} f(r_0) + \sqrt{A_0(z)} \frac{df(r_0)}{dr_0 } \right) - A(z,t) \frac{df(r_0)}{dr_0} \right) \frac{dr_0(z)}{dz}.
  \end{split}
\end{gather*}
This system of equations can be solved numerically using Richtmyer's two-step version of the Lax-Wendroff method, as done by others \cite{Olufsen2000,LeVeque1992,Kolachalama2007}. We use the code from \cite{Diem2016a}, which is an implementation of the Lax-Wendroff method in Python, to implement a model of the MCA and its first bifurcation. All simulation parameters are listed in \autoref{tab:parameter}. The MCA radius was taken from \cite{Cousins2014} and following their estimate that generally the length of an artery is $50 \cdot R_u$ lengths were calculated accordingly. The radii of the daughter vessels were estimated using a scaling factor of 0.91, also obtained from \cite{Cousins2014}. The length of the right daughter vessel was shortened slightly to avoid reflective waves cancelling one another out, as also done in \cite{Kolachalama2007}. A three-element Windkessel (3WK) model is used at the outlet whose parameters were obtained from \cite{Olufsen2002}.

\begin{figure}
\centerline{\includegraphics{figures/inlet.png}}
\caption{Inlet function for MCA simulations obtained from patient-specific measurements. The final ten peaks of the velocity time series were averaged in the Fourier space and adjusted such that its range of values aligns with those reported in \cite{Olufsen2002}. The resulting inlet function is smooth and serves as the inlet boundary condition for the MCA simulations using VaMpy \cite{Diem2016a}.\label{fig:inlet}}
\end{figure}

At the inlet flux is prescribed directly. Patient-specific flow velocity measurements were collected from the MCA of a healthy adult male using Doppler Sonography ultrasound. The velocity data was adjusted to obtain volumetric flux values within the range reported in \cite{Olufsen2002}. To obtain a smooth inlet function the final ten peaks of the time series were averaged in the Fourier space. The resulting inlet boundary condition is shown in \autoref{fig:inlet}.

\begin{figure}
\centerline{\includegraphics{figures/mca.png}}
\caption{Flux (a) and pressure (b) in the MCA. Simulations were performed using the VaMpy Python package \cite{Diem2016a} and the simulation parameters are listed in \autoref{tab:parameter}. Wall stiffness is high in small blood vessels, therefore pressure gradients in time are steep.\label{fig:mca}}
\end{figure}

MCA simulations were run for four cardiac cycles to allow the system to settle. \autoref{fig:mca} shows flow and pressure in the MCA. Because the radius of the MCA is very small its wall stiffness is high. Pressure gradients in time are therefore steep. This data provides the basis for estimating perivascular drainage through the cerebral vasculature.

\section{Perivascular Drainage through the MCA}

In the previous sections the models governing blood flow through the MCA and ISF flow through the artery wall have been introduced. Here, the results of the previous sections are coupled and used to calculate ISF flow through the BM with and without a valve mechanism to show its necessity to achieve net reverse drainage. At the same time it is shown that blood pressure driven perivascular drainage is far too slow to provide any meaningful flow, thereby disproving the popular hypothesis that arterial pulsations provide the major driving force for perivascular drainage.

\begin{figure}
\centerline{\includegraphics{figures/bm_pressure.png}}
\caption{ISF pressure inside the BM of the MCA. To obtain ISF pressure the radial stresses inside the wall were calculated from the blood pressure inside the artery. ISF pressure is directly proportional to the radial stress in the wall at the location of the BM (see Supplementary Information 2). Flow velocity was then calculated using Darcy's law.\label{fig:bm_pressure}}
\end{figure}

\begin{figure}
\centerline{\includegraphics{figures/bm_velocity.png}}
\caption{ISF velocity inside the BM of the MCA with (a) and without a valve mechanism (b) at $r = r_0 + 0.5h$. The average flow velocity with a valve mechanism is \SI{-3.54e-6}{\micro\metre\per\second} ($K_0/K_1$ = 0.01), while without a valve mechanism it is \SI{7.91e-4}{\micro\metre\per\second}. A negative value indicates net reverse flow, confirming that a valve mechanism is required to drive flow in the reverse direction of the blood. The expected magnitude of drainage is \SI{8.33}{\micro\metre\per\second} \cite{Carare2008}.\label{fig:bm_velocity}}
\end{figure}

ISF pressure inside the BM (at $r = r_0 + 0.5h$) is shown in \autoref{fig:bm_pressure}. Velocity wth and without a valve mechanism are shown in \autoref{fig:bm_velocity}. The results show that a valve mechanism is required to drive perivascular drainage in the reverse direction of the blood flow with an average flow velocity of \SI{-3.54e-6}{\micro\metre\per\second} for a ratio of $K_0/K_1 = 0.01$, while it is \SI{7.91e-4}{\micro\metre\per\second}. While the results confirm the necessity of a strong valve mechanism under pulse driven flow they also indicate that arterial pulsations are not strong enough to drive perivascular drainage. Drainage velocity is fastest without a valve mechanism (and then it is in the wrong direction), but even in that case it is four orders of magnitude slower than the expected value of \SI{8.33}{\micro\metre\per\second} \cite{Carare2008}. The valve mechanism only guarantees net reverse drainage for a permeability ratio $K_0/K_1 < 1.6e-2$.

\begin{figure}
\centerline{\includegraphics{figures/valve_test.png}}
\caption{ISF velocity at $r = r_0 + xh$. Drainage occurs in the reverse direction of the blood flow for $K_0/K_1 = \SI{1.6e-2}{}$ and is fastest at $r = r_0 + 0.1h$. The fastest net reverse drainage is \SI{-7.35e-6}{\micro\metre\per\second}, which is six orders of magnitude smaller than the expected value of \SI{-8.33}{\micro\metre\per\second}.\label{fig:valve_test}}
\end{figure}

\section{Discussion and Conclusion}

The results show that the valve mechanism as implemented in our analytical model of the BM guarantees net reverse flow inside the BM under physiological blood flow conditions. The valve mechanism is required to be strong, \ie the ratio of $K_0/K_1$ needs to be small. Additionally, the model shows that a valve mechanism is neccessary for net reverse flow to occur within the BM, confirming results from previous studies that did not include a valve mechanism and concluded the existence of some attachment mechanism \cite{Schley2006,Wang2011}. Net drainage additionally depends on the distance of the BM from the artery lumen. The closer the BM is placed to the lumen the faster net reverse drainage is, but variations remain within the same order of magnitude. This is due to the relatively large thickness of cerebral artery walls and their elastic properties. The hypothesis that the BM utilises a unidirectional valve mechanism is plausible as the general lymphatic system of the body also has one and its existence has previously been speculated \cite{Weller2010,Schley2006,Heppell2013}.

Since the development of the Schley model we have gained knowledge on the approximate velocity of the drainage and are able to perform imaging in live mice \cite{Carare2008,Schley2006,Arbel-Ornath2013}, yet this key information has not been utilised in more recent mathematical and computational studies \cite{Wang2011,Sharp2015}. From the results of Carare et al. \cite{Carare2008} it was estimated that the velocity for perivascular drainage in blood vessels of roughly \SI{10}{\micro\metre} in diameter is in the order of magnitude of \SI{8}{\micro\metre\per\second}. Arbel-Ornath et~al. \cite{Arbel-Ornath2013} measured the dye intensity of the tracers, which does not allow them to directly measure the velocity of the drainage but, from their results, a half-life period of about five minutes can be assumed for their \SI{3}{\micro\litre} injection volume. Therefore, we would expect a volumetric flux in the order of magnitude of \SI{0.001}{\cubic\milli\metre\per\second}. No correlation between ISF velocity and artery diameter was found.

The results from this study showed that the average drainage velocity of ISF is \SI{-3.54e-6}{\micro\metre\per\second} at $r = r_0 + 0.5h$ and $K_0/K_1 = 0.01$. This value indicates a net reverse flow within the BM, although it is six orders of magnitude smaller than what would be expected \cite{Carare2008}. The results therefore suggest that arterial pulsations are not powerful enough to drive perivascular drainage of ISF through the BM. This is most likely due to the very long wavelength of the arterial pulsations compared to the artery section considered, which results in very small pressure gradients along the vessel length (\autoref{fig:mca}b). It appears that other forces are neccessary to produce the pressure gradients required to push fluid through the BM and an interesting candidate for such forces is vasomotion of smooth muscle cells \cite{DiMarco2015}.

In conclusion, this study indicates that under physiological conditions arterial pulsations alone are too weak to drive perivascular drainage of ISF through the arterial BM. This result is especially interesting as arterial pulsations have been treated as the most likely candidate for the driving force of perivascular drainage. Other studies, which have considered arterial pulsations as driving mechanisms have either concluded that ISF flow in the reverse direction to the blood flow requires some form of attachment mechanism \cite{Schley2006,Wang2011} or require very specific protein movements giving rise to valve-like behaviour \cite{Sharp2015}. Our model is the first to suggest a very general valve mechanism that reliably produces net reverse drainage along the BM. While the valve model of the BM provides a phyisologically feasible mechanism to ensure net reverse drainage of ISF it is insufficient to drive the physiologically observed flow (by several orders of magnitude). This result is therefore significant as it disproves the widely accepted arterial pulsation hypothesis of perivascular drainage of \Ab from the brain \cite{Weller2009,Carare2008,Hawkes2011,Morris2014,Schley2006,Attems2011,Wang2011,Iliff2012,Asgari2015,Sharp2015,Weller2015a} and suggests that other forces have to be considered instead.

% ----------------------------------------------------------------------------
% Acknowledgements
% ----------------------------------------------------------------------------
\section*{Acknowledgements}

This work was supported by an EPSRC Doctoral Training Centre grant (EP/G03690X/1). The authors thank Tony Birch for carrying out the middle cerebral artery flow velocity measurements and Roy O. Weller for his continued advice and encouragement.

% ----------------------------------------------------------------------------
% Appendix
% ----------------------------------------------------------------------------
%\section*{Appendix}
%\appendix
%\input{mathanalysis}
%\input{airystress}

% ----------------------------------------------------------------------------
% References
% ----------------------------------------------------------------------------
\addcontentsline{toc}{section}{\bibname}

\begin{thebibliography}{10}
\bibitem{Selkoe2001}
Selkoe~DJ (2001) Alzheimer's Disease: Genes, Proteins and Therapies. {\em Physiol.~Rev.} 81(2):741--766.

\bibitem{Koffie2011}
Koffie~RM, Hyman~BT and Spires-Jones~TL (2011) Alzheimer's disease: synapses gone cold. {\em Mol.~Neurodegener.} 21(1):63.

\bibitem{Haass1992}
Haass~C et~al. (1992) Amyloid $\beta$-peptide is produced by cultured cells during normal metabolism. {\em Nature} 359(6393):322--325.

\bibitem{Weller2006}
Weller~RO, Massey~A, Kuo~YM and Roher~AE (2006) Cerebral amyloid angiopathy: accumulation of A$\beta$ in interstitial fluid drainage pathways in Alzheimer's disease. {\em Ann.~N.~Y.~Acad.~Sci.} 903(1):110--117.

\bibitem{Weller2009}
Weller~RO, Carare~RO and Boche~D (2009) Amyloid: Vascular and Parenchymal. {\em Encyclpedia of Neuroscience}, ed Squire~LR (Oxford Academic Press), pp.~355--362.

\bibitem{Weller2009a}
Weller~RO, Preston~SD, Subash~M and Carare~RO (2009) Cerebral amyloid angiopathy in the aetiology and immunotherapy of Alzheimer's disease. {\em Alzheimer's~Res.~Ther.} 1(2):6.

\bibitem{Weller2010}
Weller~RO, Galea~I, Carare~RO and Minagar~A (2010) Pathophysiology of the lymphatic drainage system of the central nervous system: Implication for pathogenesis and therapy of multiple sclerosis. {\em Pathophysiology} 17(4):295--306.

\bibitem{Carare2008}
Carare~RO et~al. (2008) Solutes, but not cells, drain from the brain parenchyma along basement membranes of capillaries and arteries: significance for cerebral amyloid angiopathy and neuroimmunology. {\em Neuropathol.~Appl.~Neurobiol.} 34(2):131--144.

\bibitem{Hawkes2011}
Hawkes~CA et~al. (2011) Perivascular drainage of solutes is impaired in the ageing mouse brain and in the presence of cerebral amyloid angiopathy. {\em Acta~Neuropathol.} 121(4):431--443.

\bibitem{Morris2014}
Morris~AWJ, Carare~RO, Schreiber~S and Hawkes~CA (2014) The cerebrovascular basement membrane: role in the clearance of $\beta$-amyloid and cerebral amyloid angiopathy. {\em Front.~Aging~Neurosci.} 6:251.

\bibitem{Schley2006}
Schley~D, Carare-Nnadi~R, Please~CP, Perry~VH and Weller~RO (2006) Mechanisms to explain the reverse perivascular transport of solutes out of the brain. {\em J.~Theor.~Biol.} 238(4):962--974.

\bibitem{Attems2011}
Attems~J, Jellinger~K, Thal~DR and Van~Nostrand~W (2011) Review: Sporadic cerebral amyloid angiopathy. {\em Neuropathol.~Appl.~Neurobiol.} 37(1):75--93.

\bibitem{Wang2011}
Wang~P and Olbricht~WL (2011) Fluid mechanics in the perivascular space. {\em J.~Theor.~Biol.} 247(1): 52--57.

\bibitem{Iliff2012}
Iliff~JJ et~al. (2012) A Paravascular Pathway Facilitates CSF Flow Through the Brain Parenchyma and the Clearance of Interstitial Solutes, Including Amyloid $\beta$. {\em Sci.~Transl.~Med.} 4(147):147ra111.

\bibitem{Asgari2015}
Asgari~M, De~Z\'elicourt~D and Kurtcuoglu~V (2015) How astrocyte networks may contribute to cerebral metabolite clearance. {\em Sci.~Rep.} 5:15024.

\bibitem{Sharp2015}
Sharp~MK, Diem~AK, Weller~RO, Carare~RO (2015) Peristalsis with Oscillating Flow Resistance: A Mechanism for Periarterial Clearance of Amyloid Beta from the Brain. {\em Ann.~Biomed.~Eng.} doi~10.1007/s10439-015-1457-6.

\bibitem{Weller2015a}
Weller~RO, Hawkes~CA, Kalaria~RN, Werring~DJ, Carare~RO (2015) White Matter Changes in Dementia: Role of Impaired Drainage of Interstitial Fluid. {\em Brain~Pathol.} 25(1):63--78.

\bibitem{Ockendon1995}
Ockendon~H and Ockendon~JR (1995) Viscous Flow, Cambridge University Press, Cambridge, UK.

\bibitem{Olufsen2000}
Olufsen~MS et~al. (2000) Numerical Simulation and Experimental Validation of Blood Flow in Arteries with Structured-Tree Outflow Conditions. {\em Ann.~Biomed.~Eng.} 28(11):1281--1299.

\bibitem{Diem2016}
Diem~AK (2016) Prediction of Perivascular Drainage of \Ab from the Brain Using Compuational Modelling: Implication for Alzheimer's Disease. {\em PhD~Thesis.}

\bibitem{LeVeque1992}
LeVeque~RJ (1992) Numerical Methods for Conservation Laws, Birkh\"auser Verlag, Basel, Switzerland.

\bibitem{Kolachalama2007}
Kolachalama~V, Bressloff~NW, Nair~PB, Shearman~CP (2007) Predictive Haemodynamics in a One-Dimensional Carotid Artery Bifurcation. Part I: Application to Stent Design. {\em IEEE~Trans.~Biomed.~Eng.} 54(5):802--812.

\bibitem{Diem2016a}
Diem~AK, Bressloff~NW (2016) VaMpy: A Python Package to Solve 1D Blood Flow Problems. {\em JORS.} Submitted.

\bibitem{Cousins2014}
Cousins~W, Gremaud~PA (2014) Impedance boundary conditions for general transient hemodynamics. {\em Int.~J.~Numer.~Method~Biomed.~Eng.}

\bibitem{Olufsen2002}
Olufsen~MS, Nadim~A, Lipsitz~LA (2002) Dynamics of cerebral blood flow regulation explained using a lumped parameter model. {\em Am.~J.~Physiol.~Regul.~Integr.~Comp.~Physiol.} 282(2):R611--R622.

\bibitem{Heppell2013}
Heppell~C, Richarson~G and Roose~T (2013) A Model for Fluid Drainage by the Lymphatic System. {\em Bull.~Math.~Biol.} 75(1):49--81.

\bibitem{Arbel-Ornath2013}
Arbel-Ornath~M et~al. (2013) Interstitial fluid drainage is impaired in ischemic stroke and Alzheimer's disease mouse models. {\em Acta~Neuropathol.} 126(3):353--364.

\bibitem{DiMarco2015}
Di~Marco~LY, Farkas~E, Martin~C, Venneri~A and Frangi~AF (2015) Is Vasomotion in Cerebral Arteries Impaired in Alzheimer's Disease? {\em J.~Alzheimers~Dis.} 46(1):35--53.
\end{thebibliography}

\end{document}

### 2017-01-11 figures/pvs_model.py

@@ -1,0 +1,55 @@
+from __future__ import division
+from matplotlib.pyplot import plot, annotate, arrow, axhline, xlim, ylim, xlabel, ylabel, xticks, yticks, savefig, show, figure, text
+from matplotlib import rc
+import numpy as np
+import os
+
+rc('font', family='sans-serif', size=9)
+rc('text', usetex=True)
+
+pi = np.pi
+end = 80*pi
+mid = end/2
+eta = np.linspace(0, end, 40*256, endpoint=True)
+Hi = 0.05*np.sin(0.05*eta)+0.99
+Ho = 0.05*np.sin(0.05*(eta+20))+1.3
+Hb = 0.05*np.sin(0.05*(eta+5))+0.5
+
+arrow_start = mid-20
+
+def cm2inch(value):
+    return value/2.54
+
+fig = figure(figsize=(cm2inch(17.2), cm2inch(8.7)))
+
+plot(eta, Hi, linewidth=2, color='k')
+plot(eta, Ho, linewidth=2, color='k')
+plot(eta, Hb, linewidth=2, color='k')
+
+axhline(y=0.99, linewidth=1, linestyle='dashed', color='k')
+axhline(y=1.3, linewidth=1, linestyle='dashed', color='k')
+axhline(y=0.5, linewidth=1, linestyle='dashed', color='k')
+
+annotate('artery wall', (mid,0.7), xycoords='data', xytext=(-22, 0), textcoords='offset points')
+annotate('ISF flow', (mid,1.15), xycoords='data', xytext=(-17,0.05), textcoords='offset points')
+annotate('artery wall', (mid,1.4), xycoords='data', xytext=(-22,0), textcoords='offset points')
+annotate('blood flow', (mid,0.2), xycoords='data', xytext=(-20,0), textcoords='offset points')
+annotate(r'$h(z,t)$', (50,1.11), xycoords='data', xytext=(5,0), textcoords='offset points')
+
+arrow(arrow_start, 0.15, 40, 0, head_width=0.05, head_length=3, color='k')
+arrow(arrow_start+40, 1.1, -40, 0, head_width=0.05, head_length=3, color='k')
+arrow(60, 1.09, 0, -0.06, head_width=2, head_length=0.02, color='k')
[…]

### 2017-01-11 paper.tex
added doctoral prize funding information
moved step function to back of section

@@
 We allow the permeability $K$ to be a function of the axial component of the ISF pressure gradient $\partial p(z,t) / \partial z = p_z(z,t)$.
-In order to obtain a net flux up the vessel, counter to the blood flow, we require this function to be increasing in $p_z$, representing the notion that it is harder to push a flow in the positive $z$-direction ($p_z<0$ and $K(p_z)$ small) than in the negative $z$-direction ($p_z>0$ and $K(p_z)$ large)
-\begin{equation}
-  K(p_z) = \begin{cases}
-    K_0 & p_z > 0\\
-    K_1 & p_z \leq 0.
-  \end{cases}
-\end{equation}
+In order to obtain a net flux up the vessel, counter to the blood flow, we require this function to be increasing in $p_z$, representing the notion that it is harder to push a flow in the positive $z$-direction ($p_z<0$ and $K(p_z)$ small) than in the negative $z$-direction ($p_z>0$ and $K(p_z)$ large).
 
@@
 $R_i(z,t)$ and $p_z$ can be obtained from a numerical artery model described in the next section.
+To model the protein valve that is being proposed here $K(p_z)$ is defined as a step function
+\begin{equation}
+  K(p_z) = \begin{cases}
+    K_0 & p_z > 0\\
+    K_1 & p_z \leq 0
+  \end{cases}
+\end{equation}
+with $K_1 > K_0$.
 
@@
 
-This work was supported by an EPSRC Doctoral Training Centre grant (EP/G03690X/1).
+This work was supported by an EPSRC Doctoral Training Centre grant (EP/G03690X/1) and EPSRC Doctoral Training Partnership grant (EP/N509747/1).
 The authors thank Tony Birch for carrying out the middle cerebral artery flow velocity measurements and Roy O.

### 2017-01-11 paper.tex
additional explanations in numerical section

@@
 \center
-\caption{Parameters used to simulate the MCA including the first bifurcation.\label{tab:parameter}}
+\caption{Geometry parameters of the MCA section modelled using VaMpy \cite{Diem2016a}.\label{tab:geometry}}
 \begin{tabular}{ccc}
@@
 $L$ & \SI{6.4}{\centi\metre} & length\\
-\midrule
-\multicolumn{3}{l}{Simulation parameters}\\
-\midrule
-$k_1$ & \SI{2.0e7}{\gram\per\centi\metre\per\square\second} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
-$k_2$ & \SI{-22.53}{\per\centi\metre} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
-$k_3$ & \SI{8.65e5}{\gram\per\centi\metre\per\square\second} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
-$R_1$ & \SI{14130}{\gram\per\centi\metre\tothe{4}\per\second} & First resistance element\\
-$R_2$ & \SI{7200}{\gram\per\centi\metre\tothe{4}\per\second} & Second resistance element\\
-$C_T$ & \SI{2.4752e-3}{\centi\metre\tothe{4}\square\second\per\gram} & Compliance element\\
-$T$ & \SI{0.85}{\second} & cardiac cycle length\\
-$t_c$ & 4 & number of cardiac cycles simulated\\
-$\Delta t$ & \SI{1e-5}{\second} & time step size\\
-$\Delta z$ & \SI{0.1}{\centi\metre} & spatial step size\\
 \bottomrule
@@
 
-Following the geometry outlined in \autoref{sec:diodemodel} arteries are modelled as axisymmetric tubes in a cylindrical coordinate system with cross-sectional area $A(z,t) = \pi R(z,t)^2$ and flux $q(z,t)$.
+In order to obtain the input functions $R_i(z,t)$ and $p_z$ for the analytical model developed in the previous section, a section of human MCA was modelled as a 1D axisymmetric tube in a cylindrical coordinate system with cross-sectional lumen area $A(z,t)$ and flux $q(z,t)$ using the Python package VaMpy \cite{Diem2016a}.
+The geometry parameters are listed in \autoref{tab:geometry}.
 Following the derivations in \cite{Olufsen2000,Diem2016} arteries are represented by the conservation law
@@
 \end{gather*}
+Elasticity of the artery wall is modelled via the quantity $f(r_0) = 4Eh/(3r_0)$. The relationship $Eh/r_0$ is based on compliance estimates
+\begin{equation}
+  \frac{Eh}{r_0} = k_1 \exp \left( k_2 r_0(z) \right) + k_3
+\end{equation}
+with $k_1, k_2, k_3$ constants \cite{Olufsen2000}.
+To derive \eqref{eq:conservation} it was necessary to make assumptions about the velocity profile of blood flow through an artery.
+Blood flow is considered pulsatile laminar and vessels are considered slightly tapered, therefore the velocity profile is assumed to be mostly flat with a thin boundary layer $\delta_b = (\nu T/(2\pi))^{0.5}$, such that $\delta_b \ll R(z,t)$ \cite{Olufsen2000}.
+
+\begin{table}
+\center
+\caption{Simulation parameters for modelling a bifurcation of the MCA using VaMpy \cite{Diem2016a}.\label{tab:simulation}}
+\begin{tabular}{ccc}
+\toprule
+Parameter & Value & Physical meaning\\
+\midrule
+$k_1$ & \SI{2.0e7}{\gram\per\centi\metre\per\square\second} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
+$k_2$ & \SI{-22.53}{\per\centi\metre} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
+$k_3$ & \SI{8.65e5}{\gram\per\centi\metre\per\square\second} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
+$\nu$ & \SI{0.046}{\square\centi\metre\per\second} & kinematic viscosity of blood\\
+$R_1$ & \SI{14130}{\gram\per\centi\metre\tothe{4}\per\second} & First resistance element\\
+$R_2$ & \SI{7200}{\gram\per\centi\metre\tothe{4}\per\second} & Second resistance element\\
+$C_T$ & \SI{2.4752e-3}{\centi\metre\tothe{4}\square\second\per\gram} & Compliance element\\
+$T$ & \SI{0.85}{\second} & cardiac cycle length\\
+$t_c$ & 4 & number of cardiac cycles simulated\\
+$\Delta t$ & \SI{1e-5}{\second} & time step size\\
+$\Delta z$ & \SI{0.1}{\centi\metre} & spatial step size\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 This system of equations can be solved numerically using Richtmyer's two-step version of the Lax-Wendroff method, as done by others \cite{Olufsen2000,LeVeque1992,Kolachalama2007}.
 We use the code from \cite{Diem2016a}, which is an implementation of the Lax-Wendroff method in Python, to implement a model of the MCA and its first bifurcation.
-All simulation parameters are listed in \autoref{tab:parameter}.
+All simulation parameters are listed in \autoref{tab:simulation}.
 The MCA radius was taken from \cite{Cousins2014} and following their estimate that generally the length of an artery is $50 \cdot R_u$ lengths were calculated accordingly.

### 2017-01-11 paper.tex
added reynolds number to numerical section

@@
 To derive \eqref{eq:conservation} it was necessary to make assumptions about the velocity profile of blood flow through an artery.
-Blood flow is considered pulsatile laminar and vessels are considered slightly tapered, therefore the velocity profile is assumed to be mostly flat with a thin boundary layer $\delta_b = (\nu T/(2\pi))^{0.5}$, such that $\delta_b \ll R(z,t)$ \cite{Olufsen2000}.
+Blood flow is considered pulsatile laminar and vessels are considered slightly tapered, therefore the velocity profile is assumed to be mostly flat with a thin boundary layer $\delta_b = (\nu T/(2\pi))^{0.5}$, such that $\delta_b \ll R(z,t)$ and $\text{Re} \approx 200$ \cite{Olufsen2000}.
 
@@
 At the same time it is shown that blood pressure driven perivascular drainage is far too slow to provide any meaningful flow, thereby disproving the popular hypothesis that arterial pulsations provide the major driving force for perivascular drainage.
+
+The numerical model of the MCA was run over four heartbeats to ensure convergence of the results. A human heartbeat lasts \SI{0.85}{\second} and the total simulation time was \SI{3.4}{\second}
 

### 2017-02-27 paper.tex
Included plots of displacement and bm width

@@
 which describes the evolution of the BM thickness and can be solved using a finite difference approximation.
-Consider an infinitesimally small volume of BM $dV = 2\pi \cdot R_i(z,t) \cdot h(z,t) \cdot dz$.
+For a detailed derivation see Supplementary Information 1.
+Another method to arrive at the same equation is derived by considering an infinitesimally small volume of BM $dV = 2\pi \cdot R_i(z,t) \cdot h(z,t) \cdot dz$.
 Then the difference between the flux into and out of the element $dV$ is equal to $\dd (dV)/ \dd t$, which can be written as
@@
 which is equal to the model equation \eqref{eq:pvs_model}.
-For a detailed derivation see Supplementary Information 1.
 The pressure $p(z,t)$ and the displacement of the BM wall $R_i(z,t)$ are, to a very good approximation, determined solely by the arterial pressure (\ie almost independent of the BM flow), see Supplementary Information 2 for more details.
@@
 \caption{Flux (a) and pressure (b) in the MCA.
-Simulations were performed using the VaMpy Python package \cite{Diem2016a} and the simulation parameters are listed in \autoref{tab:parameter}.
+Simulations were performed using the VaMpy Python package \cite{Diem2016a} and the simulation parameters are listed in Table~\ref{tab:simulation}.
 Wall stiffness is high in small blood vessels, therefore pressure gradients in time are steep.\label{fig:mca}}
@@
 At the same time it is shown that blood pressure driven perivascular drainage is far too slow to provide any meaningful flow, thereby disproving the popular hypothesis that arterial pulsations provide the major driving force for perivascular drainage.
+The numerical model of the MCA was run over four heartbeats to ensure convergence of the results.
+An averate human heartbeat lasts \SI{0.85}{\second} and the total simulation time was \SI{3.4}{\second} (see Table~\ref{tab:simulation}).
+In the results presented here only the final heartbeat is shown.
 
-The numerical model of the MCA was run over four heartbeats to ensure convergence of the results. A human heartbeat lasts \SI{0.85}{\second} and the total simulation time was \SI{3.4}{\second}
+\begin{figure}
+\centerline{\includegraphics{figures/Ri.png}}
+\caption{Displacement of the artery wall due to the pressure pulse in the MCA.
+The displacement is calculated from the linear elasticity approximation derived in Supplementary Information 2.
+The input function $R_i(z,t)$ to the BM model equation \eqref{eq:pvs_model} is obtained by adding the radius at rest $a$ (see geometry depiction in Figure~\ref{fig:pvs_model}).\label{fig:ri}}
+\end{figure}
+
+\begin{figure}
+\centerline{\includegraphics{figures/h.png}}
+\caption{BM width as calculated using the BM model equation \eqref{eq:pvs_model}. The changes in BM width are very small around the initial width of \SI{200}{\nano\metre}.\label{fig:h}}
+\end{figure}
+
+Figure~\ref{fig:ri} shows the displacement of the artery wall due to the pressure pulse.
+To caluculate the width of the BM using the model equation \eqref{eq:pvs_model}, the input function $R_i(z,t)$ is determined by adding the radius at rest $a$ to the wall displacement as depicted in Figure~\ref{fig:pvs_model}.
+Figure~\ref{fig:h} shows the width of the BM $h(z,t)$ as calculated from \eqref{eq:pvs_model}.
+The BM width only varies by $\pm \SI{0.01}{\nano\metre}$ around its initial width and thus results in very low flow rates of ISF.
 
@@
 Drainage velocity is fastest without a valve mechanism (and then it is in the wrong direction), but even in that case it is four orders of magnitude slower than the expected value of \SI{8.33}{\micro\metre\per\second} \cite{Carare2008}.
-The valve mechanism only guarantees net reverse drainage for a permeability ratio $K_0/K_1 < 1.6e-2$.
+The valve mechanism only guarantees net reverse drainage for a permeability ratio $K_0/K_1 < \SI{1.6e-2}{}$.
 

### 2017-03-10 paper.tex
Giles' feedback

@@
 \beq\label{eq:darcy}
-  \q = -K(p_z) \boldsymbol{\nabla} p
+  \q = -\frac{k}{\mu} \boldsymbol{\nabla} p
 \eeq
-where $K(p_z)$\footnote{Note that the permeability $K$ used here is the ratio of the intrinsic permeability $k$ of the ECM to the viscosity $\mu$ of ISF, that is $K = k / \mu$} is the pressure gradient dependent permeability of the ECM.
-We allow the permeability $K$ to be a function of the axial component of the ISF pressure gradient $\partial p(z,t) / \partial z = p_z(z,t)$.
-In order to obtain a net flux up the vessel, counter to the blood flow, we require this function to be increasing in $p_z$, representing the notion that it is harder to push a flow in the positive $z$-direction ($p_z<0$ and $K(p_z)$ small) than in the negative $z$-direction ($p_z>0$ and $K(p_z)$ large).
+where $k$ is the intrinsic permeability of the ECM and $\mu$ is the viscosity of ISF.
+Since ISF flows up the artery, counter to the direction of arterial pulsations, a valve-like mechanism in the ECM is required if arterial pulsations are indeed the motive force for this flow.
+This valve-like property of the ECM can be modelled by a Darcy flow \eqref{eq:darcy}, in which the intrinsic permeability is a function of the axial pressure gradient $\dd p / \dd z$ along the vessel.
+On assuming an artery aligned along the $z$-axis with net blood flow and arterial pulsations propagating the positive $z$-direction we model valve-like properties of the ECM in the BM by a pressure gradient dependent permeability
+\beq
+k = \mu K(p_z) \qquad \text{where } p_z = \frac{\dd p}{\dd z}.
+\eeq
+In order to obtain a net flux up the vessel, counter to the blood flow, we require this function to be increasing in $p_z$, representing the notion that it is harder to push a flow in the positive $z$-direction ($p_z < 0$ and $K(p_z)$ small) than in the negative $z$-direction ($p_z \geq 0$ and $K(p_z)$ large).
 
-Because of the boundary conditions used we will continue working in terms of the fluid velocity $\vel$, which is related to $\q$ by the linear relationship $\vel = \gamma \, \q$, where $\gamma$ is the porosity of the medium.
 The governing equations in cylindrical coordinates are thus
 \begin{gather}
-  u \vect{e_z} + v \vect{e_r} = - \gamma K(p_z) \left( \vect{e_z} p_z + \vect{e_r} p_r \right) \label{eq:darcy_velocity}\\
-  u_z + \frac{1}{r} \frac{\dd}{\dd r} \left( r v \right) = 0 \label{eq:continuity},
+  \boldsymbol{q} = q_1 \vect{e_z} + q_2 \vect{e_r} = - \gamma K(p_z) \left( \vect{e_z} p_z + \vect{e_r} p_r \right) \label{eq:darcy_velocity}\\
+  \frac{\dd q_1}{\dd z} + \frac{1}{r} \frac{\dd}{\dd r} \left( r q_2 \right) = 0 \label{eq:continuity},
 \end{gather}
@@
 
-Because the BM thickness is much smaller than the vessel length it is possible, on using the lubrication approximation, to derive a 1D model (in $z$) for the BM thickness $h(z,t)$; this takes the form
+Because the BM thickness is much smaller than the vessel length it is possible, on using the lubrication approximation, to derive a 1D model (in $z$) for the BM thickness $h_\textrm{bm}(z,t)$; this takes the form
 \beq
-\frac{\dd}{\dd t} \left( R_i(z,t) \cdot h(z,t) \right) = \frac{\dd}{\dd z} \left( R_i(z,t) \cdot h(z,t) \cdot K(p_{z}) \cdot p_{z} \right),\label{eq:pvs_model}
+\frac{\dd}{\dd t} \left( \gamma R_i(z,t) \cdot h_\textrm{bm}(z,t) \right) = \frac{\dd}{\dd z} \left( R_i(z,t) \cdot h_\textrm{bm}(z,t) \cdot K(p_{z}) \cdot p_{z} \right),\label{eq:pvs_model}
 \eeq
-which describes the evolution of the BM thickness and can be solved using a finite difference approximation.
+where $\gamma$ is the fluid volume fraction in the BM.
+Equation \eqref{eq:pvs_model} describes the evolution of the BM thickness and can be solved using a finite difference approximation.
 For a detailed derivation see Supplementary Information 1.
-Another method to arrive at the same equation is derived by considering an infinitesimally small volume of BM $dV = 2\pi \cdot R_i(z,t) \cdot h(z,t) \cdot dz$.
+Another method to arrive at the same equation is derived by considering an infinitesimally small volume of BM $dV = 2\pi \cdot R_i(z,t) \cdot h_\textrm{bm}(z,t) \cdot dz$.
 Then the difference between the flux into and out of the element $dV$ is equal to $\dd (dV)/ \dd t$, which can be written as
 \begin{align}
-2\pi \frac{\dd}{\dd t} \left( R_i(z,t) \cdot h(z,t) \right) &= \nonumber \\ 
-2\pi \cdot R_i(z,t) \cdot h(z,t) & \cdot q(z,t) - 2\pi \cdot R_i(z + dz,t) \cdot h(z + dz,t) \cdot q(z + dz,t),
+2\pi \frac{\dd}{\dd t} \left( \gamma R_i(z,t) \cdot h_\textrm{bm}(z,t) \right) &= \nonumber \\ 
+2\pi \cdot \gamma R_i(z,t) \cdot h_\textrm{bm}(z,t) & \cdot q(z,t) - 2\pi \cdot \gamma R_i(z + dz,t) \cdot h_\textrm{bm}(z + dz,t) \cdot q(z + dz,t),
 \end{align}
@@
   K(p_z) = \begin{cases}
-    K_0 & p_z > 0\\
-    K_1 & p_z \leq 0
+    K_0 & p_z < 0\\
+    K_1 & p_z \geq 0,
   \end{cases}
 \end{equation}
-with $K_1 > K_0$.
+chosen such that $K_1 > K_0 > 0$ so that the BM is more permeable to flow of ISF in the negative $z$-direction than in the positive one.
 
@@
 The geometry parameters are listed in \autoref{tab:geometry}.
-Following the derivations in \cite{Olufsen2000,Diem2016} arteries are represented by the conservation law
+Following the derivations in \cite{Olufsen2000,Diem2016} blood flow through arteries is represented by the conservation law
 \begin{equation}
@@
 \end{gather*}
-Elasticity of the artery wall is modelled via the quantity $f(r_0) = 4Eh/(3r_0)$. The relationship $Eh/r_0$ is based on compliance estimates
+The equation is solved for cross-sectional area $A(z,t)$ of the vessel lumen and flux $q(z,t)$ where $R(z,t) = q(z,t)/A(z,t)$ is the lumen radius, $\delta_b$ is the boundary layer thickness and $\text{Re}$ is the Reynolds number.
+Elasticity of the artery wall is modelled via the quantity $f(r_0) = 4Eh/(3r_0)$, where $E$ is the Young's modulus and $h$ is the wall thickness.
+The relationship $Eh/r_0$ is based on compliance estimates
 \begin{equation}
@@
 \end{equation}
-with $k_1, k_2, k_3$ constants \cite{Olufsen2000}.
+with $k_1, k_2, k_3$ constants \cite{Olufsen2000}. Blood pressure $p(z,t)$ is related to $A(z,t)$ and $f(r_0)$ via the state equation
+\beq
+  p(z,t) - p_0 = f(r_0) \left( 1 - \frac{A_0(z)}{A(z,t)} \right),
+\eeq
+where $p_0$ is the diastolic pressure and $A_0(z)$ is the lumen radius at rest.
+Blood pressure provides the basis for determining ISF pressure inside the BM via the calculations in Supplementary Information 2.
+
 To derive \eqref{eq:conservation} it was necessary to make assumptions about the velocity profile of blood flow through an artery.
@@
 \centerline{\includegraphics{figures/inlet.png}}
-\caption{Inlet function for MCA simulations obtained from patient-specific measurements.
+\caption{Inlet function for MCA simulations obtained from patient-specific measurements used to represent one period of \SI{0.85}{\second} of the pulse wave.
 The final ten peaks of the velocity time series were averaged in the Fourier space and adjusted such that its range of values aligns with those reported in \cite{Olufsen2002}.
@@
 
-At the inlet flux is prescribed directly.
+At the inlet the flux is prescribed directly.
 Patient-specific flow velocity measurements were collected from the MCA of a healthy adult male using Doppler Sonography ultrasound.
@@
 \begin{figure}
-\centerline{\includegraphics{figures/h.png}}
-\caption{BM width as calculated using the BM model equation \eqref{eq:pvs_model}. The changes in BM width are very small around the initial width of \SI{200}{\nano\metre}.\label{fig:h}}
+\centerline{\includegraphics{figures/h_dh.png}}
+\caption{BM width $h_\textrm{bm}(z,t)$ at several time points during the cardiac cycle (a) and its gradient along the $z$-axis at the same time points (b).
+Pressure pulse driven variations of the BM width are small compared to the initial width of \SI{200}{\nano\metre} and thus flow through the BM is slow.\label{fig:h}}
 \end{figure}
@@
 Figure~\ref{fig:ri} shows the displacement of the artery wall due to the pressure pulse.
-To caluculate the width of the BM using the model equation \eqref{eq:pvs_model}, the input function $R_i(z,t)$ is determined by adding the radius at rest $a$ to the wall displacement as depicted in Figure~\ref{fig:pvs_model}.
-Figure~\ref{fig:h} shows the width of the BM $h(z,t)$ as calculated from \eqref{eq:pvs_model}.
-The BM width only varies by $\pm \SI{0.01}{\nano\metre}$ around its initial width and thus results in very low flow rates of ISF.
+To caluculate the width of the BM using the model equation \eqref{eq:pvs_model}, the input function $R_i(z,t)$ is the sum of the radius at rest and the wall displacement, as depicted in Figure~\ref{fig:pvs_model}.
+Figure~\ref{fig:h}a shows the width of the BM $h_\textrm{bm}(z,t)$ as calculated from \eqref{eq:pvs_model} over the length of the artery at a number of time points, while Figure~\ref{fig:h}b shows $\partial h_\textrm{bm}(z,t)/ \partial z$ for the same time points.
+The BM width only varies by approximately $\pm \SI{0.01}{\nano\metre}$ around its initial width and thus results in very low flow rates of ISF.
 
@@
 \begin{figure}
-\centerline{\includegraphics{figures/bm_velocity.png}}
-\caption{ISF velocity inside the BM of the MCA with (a) and without a valve mechanism (b) at $r = r_0 + 0.5h$.
-The average flow velocity with a valve mechanism is \SI{-3.54e-6}{\micro\metre\per\second} ($K_0/K_1$ = 0.01), while without a valve mechanism it is \SI{7.91e-4}{\micro\metre\per\second}.
-A negative value indicates net reverse flow, confirming that a valve mechanism is required to drive flow in the reverse direction of the blood.
-The expected magnitude of drainage is \SI{8.33}{\micro\metre\per\second} \cite{Carare2008}.\label{fig:bm_velocity}}
-\end{figure}
-
-ISF pressure inside the BM (at $r = r_0 + 0.5h$) is shown in \autoref{fig:bm_pressure}.
-Velocity wth and without a valve mechanism are shown in \autoref{fig:bm_velocity}.
-The results show that a valve mechanism is required to drive perivascular drainage in the reverse direction of the blood flow with an average flow velocity of \SI{-3.54e-6}{\micro\metre\per\second} for a ratio of $K_0/K_1 = 0.01$, while it is \SI{7.91e-4}{\micro\metre\per\second}.
-While the results confirm the necessity of a strong valve mechanism under pulse driven flow they also indicate that arterial pulsations are not strong enough to drive perivascular drainage.
-Drainage velocity is fastest without a valve mechanism (and then it is in the wrong direction), but even in that case it is four orders of magnitude slower than the expected value of \SI{8.33}{\micro\metre\per\second} \cite{Carare2008}.
-The valve mechanism only guarantees net reverse drainage for a permeability ratio $K_0/K_1 < \SI{1.6e-2}{}$.
-
-\begin{figure}
 \centerline{\includegraphics{figures/valve_test.png}}
@@
 \end{figure}
+
+ISF pressure inside the BM (at $r = r_0 + 0.5h$) is shown in \autoref{fig:bm_pressure}.
+The results show that a valve mechanism is required to drive perivascular drainage in the reverse direction of the blood flow with an average flow velocity of \SI{-3.54e-6}{\micro\metre\per\second} for a ratio of $K_0/K_1 = 0.01$, while it is \SI{7.91e-4}{\micro\metre\per\second}.
+While the results confirm the necessity of a strong valve mechanism under pulse driven flow they also indicate that arterial pulsations are not strong enough to drive perivascular drainage.
+Drainage velocity is fastest without a valve mechanism (and then it is in the wrong direction), but even in that case it is four orders of magnitude slower than the expected value of \SI{8.33}{\micro\metre\per\second} \cite{Carare2008}.
+The valve mechanism only guarantees net reverse drainage for a permeability ratio $K_0/K_1 < \SI{1.6e-2}{}$, where $K_1$ is chosen such that $k = \SI{1e-10}{\square\centi\metre}$ is consistent with values for other tissues in the body and $\mu = \SI{1.5e-3}{\pascal\second}$ \cite{Heppell2013}.
 
@@
 This work was supported by an EPSRC Doctoral Training Centre grant (EP/G03690X/1) and EPSRC Doctoral Training Partnership grant (EP/N509747/1).
-The authors thank Tony Birch for carrying out the middle cerebral artery flow velocity measurements and Roy O.
+The authors thank Tony Birch for carrying out the middle cerebral artery flow velocity measurements, Roxana Aldea for our discussions on the appropriate equations for artery wall elasticity and Roy O.
 Weller for his continued advice and encouragement.

### 2017-03-10 figures/pvs_model.py

@@ -34,7 +34,7 @@ annotate('artery wall', (mid,0.7), xycoords='data', xytext=(-22, 0), textcoords=
 annotate('ISF flow', (mid,1.15), xycoords='data', xytext=(-17,0.05), textcoords='offset points')
 annotate('artery wall', (mid,1.4), xycoords='data', xytext=(-22,0), textcoords='offset points')
 annotate('blood flow', (mid,0.2), xycoords='data', xytext=(-20,0), textcoords='offset points')
-annotate(r'$h(z,t)$', (50,1.11), xycoords='data', xytext=(5,0), textcoords='offset points')
+annotate(r'$h_\textrm{bm}(z,t)$', (50,1.11), xycoords='data', xytext=(5,0), textcoords='offset points')
 
 arrow(arrow_start, 0.15, 40, 0, head_width=0.05, head_length=3, color='k')
 arrow(arrow_start+40, 1.1, -40, 0, head_width=0.05, head_length=3, color='k')

### 2017-03-13 paper.tex
experimental results added

@@
 
+\DeclareSIUnit\Molar{\textsc{M}}
+
 \begin{document}
@@
 \author[3]{Roxana O. Carare}
+\author[3]{Matthew MacGregor Sharp}
 \author[4]{Giles Richardson}
@@
 Therefore arterial pulsations are concluded to be insufficient to drive perivascular drainage along the BM of cerebral arteries.
+The simulation results are confirmed by an experimental assessment of the effect of pulse modulation in a mouse model, which showed no significant difference in perivascular drainage compare to control mice.
 
-\section{Porous medium model of lymphatic drainage in the BM}
+\section{Numerical modelling of perivascular drainage of \Ab}
+
+First, the velocity of perivascular drainage in a pressure-pulse driven scenario is evaluated using computational modelling.
+An analytical model of the BM was developed, which captures the preference of flow in the reverse direction to the blood flow.
+Pressure inside the BM is evaluated using a numerical model of blood pressure inside the artery and a stress analysis in the artery wall.
+These were then used to evaulate the velocity of perivascular drainage through the MCA.
+
+\subsection{Porous medium model of lymphatic drainage through the BM}
 \label{sec:diodemodel}
@@
 
-\section{Numerical Model of a Straight Arterial Section}
+\subsection{Numerical Model of a Straight Arterial Section}
 
@@
 
-\section{Perivascular Drainage through the MCA}
+\subsection{Perivascular Drainage through the MCA}
 
@@
 
+\section{Experimental assessment of intramural perivascular drainage of \Ab}
+
+In addition to performing simulations to assess the feasibility of blood-pressure driven arterial pulsations providing the main driving mechanism for perivascular drainage through the BM of the cerebral vasculature, mouse model experiments were performed to confirm the results \textit{in vivo}.
+
+\subsection{Intracerebral injections}
+
+All mice were kept on a standard 12-hr light/dark cycle and allowed food and water \textit{ad libitum}.
+All experiments were carried out in accordance with animal care guidelines stipulated by the Animal Care and Use Committee at the University of Southampton and the Home Office (PPL 20/2095).
+
+To minimise the effects of anaesthesia on mouse cardiovascular function, isoflurane was chosen as the anaesthetizing agent.
+To reduce arterial pulsations, the beta-blocker Atenolol (Ate, \SI{10}{\milli\gram\per\kilo\gram}) was administered intraperitoneally (IP).
+Mouse vital parameters were monitored throughout the procedure using a MouseOX Plus and an infrared mouse thigh sensor (STARR Life Sciences, Holliston, MA, USA).
+Thermal regulation was maintained using a heat pad and rectal probe (BASi, West Lafayette, IN, USA).
+All intracerebral injections were performed using a \SIrange{1}{5}{\micro\litre} Hirschmann microcapillary pipette (Sigma UK) with a tip adjusted to a diameter of $< \SI{50}{\micro\metre}$ using a Sutter P97 Flaming Brown Pipette puller.
+
+Ten week old wild-type C57B16 mice ($n = 4$) were anaesthetized with isoflurane, placed prone on a stereotaxic instrument and secured with head adaptors.
+Baseline measurements of oxygen saturation, heart rate, breath rate and pulse distention were taken before administering Ate.
+A volume of \SI{0.5}{\micro\litre} of \Abs HiLyte Fluor 555 (Cambridge Bioscience, UK) was stereotaxically injected into the left hippocampus at a rate of \SI{0.2}{\micro\litre\per\minute} (coordinates from Bregma: $\textrm{AP} = \SI{-1.9}{\milli\metre}$, $\textrm{ML} = \SI{1.5}{\milli\metre}$ and $\textrm{DV} = \SI{-1.7}{\milli\metre}$) \SI{10}{\minute} post Ate injection.
+Injection pipettes were left \textit{in situ} for \SI{2}{\minute} to prevent reflux.
+Mice were euthanized \SI{5}{\minute} after withdrawal of the injection pipette through overdose with pento-barbital (\SI{200}{\milli\gram\per\kilo\gram}).
+Mice were then intracardially perfused with \SI{0.01}{\Molar} phosphate buffered saline (PBS), pH~7.2 followed by \SI{4}{\percent} paraformaldehyde in \SI{0.01}{\Molar} PBS, pH~7.2.
+Brains were removed, post fixed overnight in fresh fixative and then processed for immunohistochemistry.
+For controls, ten week old wild-type C57B16 ($n = 3$) were processed following the same method but without administration of Ate.
+
+\subsection{Immunohistochemistry}
+
+Brains were cut into \SI{20}{\micro\metre} coronal sections using a freezing microtome (Leica) and the injection site located by examining on a Zeiss Axioskop 2 fitted with a rhodamine filter.
+Sections of \SI{400}{\micro\metre} and \SI{800}{\micro\metre} thickness both posterior and anterior to the injection site were selected for immunohistochemistry.
+Sections were blocked in \SI{15}{\percent} goat serum (Sigma 9023) for \SI{1}{\hour} at room temperature (RT).
+They were then incubated in primary antibodies diluted in \SI{0.01}{\Molar} phosphate buffered saline \SI{0.1}{\percent} triton x100 (PBSt).
+Primary antibodies comprised rabbit anti-collagen IV (AbCam, ab6586) 1/400 and mouse anti-smooth muscle actin (SMA) FITC conjugated (Sigma F3777) 1/200.
+Sections were incubated overnight at \SI{4}{\celsius} in a moist chamber.
+Following washing 3x\SI{10}{\minute} \SI{0.01}{\Molar} PBS sections were incubated in goat anti-rabbit Alexa Fluor$\textregistered$ 633 (ThermoFisher Scientific A-21070) 1/200 PBSt.
+
+\subsection{Image analysis}
+
+Images of the whole hippocampus were captured on a Leica SP8 confocal microscope fitted with LASX software in the tile scanning mode with sequential imaging.
+Once captured, images were exported as TIFF files and uploaded into Adobe Photoshop CS6.
+An area of hippocampus was drawn extending transversely from CA3 pyramidal cells to the apex of the blades of the granule cell layer and from the ventricle to \textit{stratum oriens} in the superior/inferior axis.
+\Ab containing vessels were defined according to size and presence of SMA: capillaries had a diameter of $\leq \SI{10}{\micro\metre}$, veins had no SMA and a diameter of $> \SI{10}{\micro\metre}$.
+The number of vessels was expressed per \SI{}{\square\milli\metre} of hippocampus.
+
+Statistical analysis was performed by an independent t-test, with \SI{95}{\percent} set as confidence interval.
+
+\subsection{Results}
+
+\begin{figure}
+  \centerline{\includegraphics{figures/confocal.jpg}}
+  \caption{Caption\label{fig:confocal}}
+\end{figure}
+
+\begin{table}
+  \center
+  \caption{Caption\label{tab:caption}}
+  \begin{tabular}{lcccccc}
+    \toprule
+     & \multicolumn{3}{c}{Atenolol} & \multicolumn{3}{c}{Saline}\\
+    \midrule
+    arteries    & 6.19 & 0.66 & 4 & 7.48  & 2.4  & 3\\
+    veins       & 0.79 & 0.3  & 4 & 0.3   & 0.15 & 3\\
+    capillaries & 0.52 & 0.14 & 4 & 0.6   & 0.3  & 3\\
+    \midrule
+    arteries    & 6.79 & 1.4  & 4 & 15.03 & 5.26 & 3\\
+    veins       & 0.62 & 0.05 & 4 & 1.8   & 1.08 & 3\\
+    capillaries & 0.17 & 0.1  & 4 & 2.34  & 1.55 & 3\\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\begin{figure}
+  \centerline{\includegraphics{figures/graph.png}}
+  \caption{Caption\label{fig:graph}}
+\end{figure}
+
+Within \SI{5}{\minute} of injections into the hippocampus, fluorescent \Ab was observed diffusely in the hippocampus and co-localizing with collagen IV in the walls of capillaries and arteries, with very few veins involved in both atenolol-treated and control mice (Figure~\ref{fig:caption} and Table~\ref{tab:confocal}).
+Application of an independent t-test to the data revealed no significant differences between the saline and atenolol treated mice either anterior or posterior to the injection site (Figure~\ref{fig:graph}).
+
 \section{Discussion and Conclusion}
@@
 The valve mechanism is required to be strong, \ie the ratio of $K_0/K_1$ needs to be small.
-Additionally, the model shows that a valve mechanism is neccessary for net reverse flow to occur within the BM, confirming results from previous studies that did not include a valve mechanism and concluded the existence of some attachment mechanism \cite{Schley2006,Wang2011}.
+Additionally,the model shows that a valve mechanism is neccessary for net reverse flow to occur within the BM, confirming results from previous studies that did not include a valve mechanism and concluded the existence of some attachment mechanism \cite{Schley2006,Wang2011}.
 Net drainage additionally depends on the distance of the BM from the artery lumen.

### 2017-07-18 paper.tex
frontiers submission

--- 2017-03-10 paper.tex
@@
-\documentclass[a4paper,titlepage]{scrartcl}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{amsmath,amssymb,amsthm}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
+% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.                                                 %
+%                                                                                                                                                         %
+% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%% Version 3.1 Generated 2015/22/05 %%%
+%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
+%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
+
+%\documentclass{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
+%\documentclass{frontiersHLTH} % for Health articles
+\documentclass{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles
+
+%\setcitestyle{square}
+\usepackage{url,hyperref,lineno,microtype}
+\usepackage[onehalfspacing]{setspace}
+\usepackage{textgreek}
+\usepackage{xspace}
 \usepackage{siunitx}
-\usepackage{booktabs}
-\usepackage[format=hang,margin=1cm]{caption}
-\usepackage{rotating}
-\usepackage{subcaption}
-\usepackage{authblk}
-\usepackage{color}
-\usepackage{xspace}
-\usepackage{textgreek}
+\linenumbers
+
+\DeclareSIUnit\Molar{\textsc{M}}
+\newcommand{\Ab}{A\textbeta\xspace}
+\newcommand{\ie}{i.~e.\xspace}
+\newcommand{\q}{\boldsymbol{q}}
+\newcommand{\dd}{\partial}
+\newcommand{\vect}[1]{\boldsymbol{#1}}
+\newcommand{\Abs}{A\textbeta\textsubscript{1-40}\xspace}
+
+% Leave a blank line between paragraphs instead of using \\
+
+\def\keyFont{\fontsize{8}{11}\helveticabold }
+\def\firstAuthorLast{Diem {et~al.}} %use et al only if is more than 1 author
+\def\Authors{Alexandra K. Diem\,$^{1,*}$, Neil W. Bressloff\,$^{1}$, Roxana O. Carare\,$^{2}$, Matthew MacGregor Sharp\,$^{2}$ and Giles Richardson\,$^3$}
+% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
+% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
+\def\Address{$^{1}$Computational Engineering and Design, Faculty of Engineering \& the Environment, University of Southampton, Southampton SO16 7QF, UK \\
+$^{2}$Clinical Neurosciences, Faculty of Medicine, University of Southampton, Southampton General Hospital, Southampton SO16 6YD, UK\\
+$^{3}$Mathematical Sciences, Faculty of Social, Human and Mathematical Sciences, University of Southampton, Southampton SO17 1BJ, UK}
+% The Corresponding Author should be marked with an asterisk
+% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
+\def\corrAuthor{A.~K.~Diem}
+\def\corrAddress{Computational Engineering and Design, Faculty of Engineering \& the Environment, University of Southampton, Southampton SO16 7QF, UK}
+\def\corrEmail{A.K.Diem@soton.ac.uk}
 
 \begin{document}
+\onecolumn
+\firstpage{1}
 
-\input{definitions}
+\title[Arterial pulsations cannot drive periarterial drainage]{Arterial pulsations cannot drive intramural periarterial drainage: Significance for Alzheimer's disease}
 
-\sisetup{range-phrase=--}
-\sisetup{range-units=single}
+\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
+\address{} %This field will be automatically populated
+\correspondance{} %This field will be automatically populated
 
-% ----------------------------------------------------------------------------
-% Titlepage Parameters
-% ----------------------------------------------------------------------------
-
-\title{Arterial pulsations are not powerful enough to drive perivascular drainage of solutes from the brain}
-\author[1,2]{Alexandra K. Diem}
-\author[2]{Neil W. Bressloff}
-\author[3]{Roxana O. Carare}
-\author[4]{Giles Richardson}
-\affil[1]{Institute for Complex Systems Simulation, University of Southampton, Southampton SO17 1BJ, UK}
-\affil[2]{Computational Engineering and Design, Faculty of Engineering \& the Environment, University of Southampton, Southampton SO17 1BJ, UK}
-\affil[3]{Faculty of Medicine, University of Southampton, Southampton, UK}
-\affil[4]{Faculty of Mathematical Sciences, University of Southampton, Southampton, UK}
+\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
+%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% The sections below are for reference only.
+%%%
+%%% For Original Research Articles, Clinical Trial Articles, and Technology Reports the section headings should be those appropriate for your field and the research itself.
+It is recommended to organize your manuscript in the
+%%% following sections or their equivalents for your field:
+%%% Abstract, Introduction, Material and Methods, Results, and Discussion.
+%%% Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
+%%%
+%%%For information about Clinical Trial Registration, please go to http://www.frontiersin.org/about/AuthorGuidelines#ClinicalTrialRegistration
+%%%
+%%% For Clinical Case Studies the following sections are mandatory: Abstract, Introduction, Background, Discussion, and Concluding Remarks.
+%%%
+%%% For all other article types there are no mandatory sections.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{abstract}
-\textbf{Summary.} Alzheimer's Disease (AD) is the most common form of dementia and to date there is no preventive medication.
-Its development is closely related to the accumulation of the neuronal metabolite \Ab, which raises the question of how metabolites are eliminated from the brain.
-Unlike the rest of the body the brain does not have a lymphatic system and therefore has to eliminate metabolites differently.
-A connection between the pathology of AD and cerebrovascular diseases has been suggested and confirmed in various mice studies, which has lead to the assumption that interstitial fluid in the basement membranes of cerebral arteries provide the pathways for the perivascular drainage of \Ab.
-However, perivascular drainage occurs in the reverse direction to arterial blood flow, an effect that lacks an effective explanation.
-In this study we tested the well-accepted hypothesis that arterial pulsations drive perivascular drainage and additionally that a valve mechanism exists which ensure net reverse drainage.
-This was examined in terms of a  mathematical model of the  drainage mechanism.
-It is shown that a valve mechanism (such as directional permeability of the perivascular space) is necessary to achieve a net reverse flow.
-However, it is also shown that arterial pulsations are not strong enough to produce drainage velocities comparable to experimental observations.
-% 209 words
 
-Word count: 209
+%%% Leave the Abstract empty if your article falls under any of the following categories: Editorial Book Review, Commentary, Field Grand Challenge, Opinion or specialty Grand Challenge.
+  \section{
+    Alzheimer's Disease (AD) is the most common form of dementia and to date there is no cure or efficient prophylaxis.
+The cognitive decline correlates with the accumulation of amyloid-\textbeta\xspace (\Ab) in the walls of capillaries and arteries.
+Our group has demonstrated that interstitial fluid and \Ab are eliminated from the brain along the basement membranes of capillaries and arteries, the intramural periarterial drainage (IPAD) pathway.
+With advancing age and arteriosclerosis, the stiffness of arterial walls, this pathway fails in its function and \Ab accumulates in the walls of arteries.
+In this study we tested the hypothesis that arterial pulsations drive IPAD and that a valve mechanism ensures the net drainage in a direction opposite to that of the blood flow.
+This hypothesis was tested using a mathematical model of the  drainage mechanism.
+We demonstrate firstly that arterial pulsations are not strong enough to produce drainage velocities comparable to experimental observations.
+Secondly, we demonstrate that a valve mechanism such as directional permeability of the IPAD pathway is necessary to achieve a net reverse flow.
+The mathematical simulation results are confirmed by assessing the pattern of IPAD in mice using pulse modulators, showing no significant alteration of IPAD.
+Our results indicate that forces other than the cardiac pulsations are responsible for efficient IPAD.
+  }
+%As a primary goal, the abstract should render the general significance and conceptual advance of the work clearly accessible to a broad readership. References should not be cited in the abstract.
+%For full guidelines regarding your manuscript please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines}{Author Guidelines} \\ or \textbf{Table \ref{Tab:01}} for a summary according to article type.
 
-\vspace{1cm}
-
-\textbf{Keywords:} Alzheimer's disease, brain lymphatics, cerebrovascular flow
+\tiny
+ \keyFont{ \section{Keywords:} Alzheimer's disease; cerebral blood flow; perivascular drainage; intramural periarterial drainage; cerebral lymphatics } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
 \end{abstract}
 
-% ----------------------------------------------------------------------------
-% Paper Body
-% ----------------------------------------------------------------------------
+\section{Introduction}
 
-% Introduction
-
-Dementia, a term often used synonymously with Alzheimer's disease (AD), is a symptom of a variety of diseases affecting the human brain and does not constitute a part of normal ageing.
+Alzheimer's disease (AD) is the commonest form of dementia.
 Although it has been studied for over 100 years \cite{Selkoe2001}, to date there is no cure as the processes relevant to its onset and progression have not yet been fully understood.
-The pathology of AD comprises different characteristics, some of which are degeneration and death of neurons, synapse loss, neuroinflammation and the accumulation of waste product deposits \cite{Koffie2011}.
-The protein amyloid-$\beta$ (\Ab) is produced as part of the normal metabolism of healthy brains, but it plays a major role in the development of AD \cite{Haass1992}.
-Its accumulation in the walls of cerebral blood vessels as cerebral amyloid angiopathy (CAA) is closely associated with the onset and progression of AD \cite{Weller2006, Weller2009}.
-CAA strongly correlates with the progression of AD and the degree of dementia as it appears in \SIrange{90}{96}{\percent} of AD patients \cite{Weller2009a}.
+The pathology of AD comprises the degeneration and death of neurons, synapse loss, neuroinflammation and the intracellular as well as extracellular accumulation of proteins \cite{Koffie2011}.
+The protein amyloid-\textbeta\xspace (\Ab) is produced as part of the normal metabolism of healthy brains, but it plays a major role in the development of AD \cite{Haass1992}.
+Its accumulation in the walls of cerebral blood vessels as cerebral amyloid angiopathy (CAA) is closely associated with the onset and progression of AD.
+CAA strongly correlates with the progression of AD and the degree of dementia as it appears in \SIrange{90}{96}{\percent} of AD patients \cite{Weller2006, Weller2009}.
 
@@
 While the rest of the body possesses a dedicated network of lymphatic vessels these do not exist in the brain \cite{Weller2010}.
-There is ample evidence that the basement membranes (BM) in the walls of cerebral arteries provide the clearance pathway for interstitial fluid (ISF) towards the surface of the brain, also referred to as perivascular drainage \cite{Weller2010,Carare2008,Hawkes2011,Morris2014}.
-Because it provides the pathways for perivascular drainage, the BM is also referred to as perivascular space (PVS) in this context.
-It is important not to confuse this PVS with the Virchow-Robin spaces, which are also often referred to as perivascular (or ``paravascular'') spaces.
-However, Virchow-Robin spaces are virtual spaces and have no role in \Ab drainage.
-Additionally, the BM does not actually constitute a space, but rather a fluid-filled protein matrix.
+There is ample evidence that the basement membranes (BM) in the walls of cerebral arteries provide the clearance pathway for interstitial fluid (ISF) towards the surface of the brain, also referred to as intramural periarterial drainage (IPAD) \cite{Weller2010,Carare2008,Hawkes2011,Morris2014}.
+It is important not to confuse the IPAD pathway with the Virchow-Robin spaces, which are often referred to as perivascular (or ``paravascular'') spaces, possibly artefactual, located between the glia limitans and the arterial wall.
+Virchow-Robin spaces are virtual spaces and have no role in \Ab drainage.
+The BM forming the IPAD does not actually constitute a space, but rather a fluid-filled protein matrix.
 Our understanding of the mechanism that drives perivascular drainage through the BM is still very limited.
-The evidence indicates that ISF enters the BM at the capillary level and flows towards the lymph nodes in the neck via cerebral arteries, but not veins.
-This suggests that ISF flow is driven in the reverse direction of the blood flow, a phenomenon that does not seem very plausible at first sight \cite{Carare2008}.
+The evidence indicates that ISF enters the BM at the capillary level and flows towards the lymph nodes in the neck via the BM cerebral arteries.
+This suggests that the counterintuitive phenomenon of ISF flow occurs in the reverse direction to blood flow \cite{Carare2008}.
 
@@
 In this paper we tested this hypothesis by using a multi-scale modelling approach combining analytical and numerical methods.
-The BM is modelled as a porous medium using Darcy's law and a lubrication approximation.
+The BM is modelled as a porous medium using a lubrication approximation.
 Flow inside the BM is driven by both the ISF pressure gradient and the deformation of the arterial wall, which are obtained from a numerical model of the middle cerebral artery (MCA).
-To guarantee reverse drainage (\ie in the opposite direction of the blood flow) a valve mechanism is implemented as a pressure gradient dependent permeablility.
-It is shown that this valve mechanism is neccesary to achieve net reverse drainage.
-The results show that this valve mechanism successfully ensures net reverse drainage inside the BM and is neccesary.
-However, due to the long wavelength of arterial pulsations the pressure gradients inside the BM were found to be too small to drive fluid flow at the expected velocities.
-The resulting velocity of the model was approximately two orders of magnitude smaller than values obtained experimentally \cite{Carare2008}.
-Therefore arterial pulsations are concluded to be insufficient to drive perivascular drainage along the BM of cerebral arteries.
+To give IPAD in the reverse direction to the blood flow, a valve mechanism is required; here it is modelled by a pressure gradient dependent permeability.
+It is shown that this valve mechanism is necessary to achieve net reverse drainage.
+However, due to the long wavelength of arterial pulsations the pressure gradients inside the BM are too small to drive fluid flow at the expected velocities.
+The resulting velocity of the model was at least four orders of magnitude smaller than values obtained experimentally \cite{Carare2008}.
+Therefore, arterial pulsations are concluded to be insufficient to drive perivascular drainage along the BM of cerebral arteries.
+The simulation results are confirmed by an experimental assessment of the effect of pulse modulation in a mouse model, which showed no significant difference in IPAD compare to control mice.
 
-\section{Porous medium model of lymphatic drainage in the BM}
+%\begin{methods}
+\section{Numerical modelling of perivascular drainage of \Ab}
+
+First, IPAD under a pressure-pulse driven scenario is evaluated using computational modelling.
+An analytical model of the BM was developed, which captures the preference of flow in the reverse direction to the blood flow.
+Pressure inside the BM is evaluated using a numerical model of blood pressure inside the artery and a stress analysis in the artery wall.
+These were then used to evaluate the velocity of perivascular drainage through the MCA.
+
+\subsection{Porous medium model of lymphatic drainage through the BM}
 \label{sec:diodemodel}
@@
 Such a pressure gradient dependent permeability could, for instance, model proteins in the BM protein network that have structural properties that allow them to bend easily only in one direction, thereby acting as a valve, and thus provide a higher resistance to flow in the direction of the blood flow.
-A similar idea has been investigated in \cite{Sharp2015}, where proteins were explicitly modelled as cillia.
+A similar idea has been investigated in \cite{Sharp2015}, where proteins were explicitly modelled as cilia.
 However, here we model the active component of the BM generically in this model and thus allow for alternative explanations of the phenomenon.
@@
 Darcy's law states that the flux $\q$ is related to the fluid pressure gradient $\boldsymbol{\nabla} p$ by
-\beq\label{eq:darcy}
+\begin{equation}\label{eq:darcy}
   \q = -\frac{k}{\mu} \boldsymbol{\nabla} p
-\eeq
+\end{equation}
 where $k$ is the intrinsic permeability of the ECM and $\mu$ is the viscosity of ISF.
@@
 On assuming an artery aligned along the $z$-axis with net blood flow and arterial pulsations propagating the positive $z$-direction we model valve-like properties of the ECM in the BM by a pressure gradient dependent permeability
-\beq
+\begin{equation}
 k = \mu K(p_z) \qquad \text{where } p_z = \frac{\dd p}{\dd z}.
-\eeq
+\end{equation}
 In order to obtain a net flux up the vessel, counter to the blood flow, we require this function to be increasing in $p_z$, representing the notion that it is harder to push a flow in the positive $z$-direction ($p_z < 0$ and $K(p_z)$ small) than in the negative $z$-direction ($p_z \geq 0$ and $K(p_z)$ large).
@@
 \begin{gather}
-  \boldsymbol{q} = q_1 \vect{e_z} + q_2 \vect{e_r} = - \gamma K(p_z) \left( \vect{e_z} p_z + \vect{e_r} p_r \right) \label{eq:darcy_velocity}\\
+  \q = q_1 \vect{e_z} + q_2 \vect{e_r} = - K(p_z) \left( \vect{e_z} p_z + \vect{e_r} p_r \right) \label{eq:darcy_velocity}\\
   \frac{\dd q_1}{\dd z} + \frac{1}{r} \frac{\dd}{\dd r} \left( r q_2 \right) = 0 \label{eq:continuity},
 \end{gather}
-where $\vect{e_z}$ and $\vect{e_r}$ describe the unit vectors in the $z$ and $r$ direction. The continuity of velocity is given in \eqref{eq:continuity}.
-
-\begin{figure}
-\centerline{\includegraphics{figures/pvs_model.pdf}}
-\caption{Geometry of the perivascular drainage pathways model.
-The length $L$ of a typical artery in the brain can be assumed to be much greater than the width $W$ of the perivascular drainage pathways.
-Therefore we can approximate the flow as a lubrication model.
-The pulsation of the capillary is described by the function $h(z,t)$.\label{fig:pvs_model}}
-\end{figure}
+where $\vect{e_z}$ and $\vect{e_r}$ describe the unit vectors in the $z$ and $r$ direction. Incompressibility of the fluid is represented by \eqref{eq:continuity}.
 
@@
 Assuming that the flows through the surfaces of the BM are negligible the following kinematic boundary condition applies
-\beq \label{eq:kinematic}  
+\begin{equation} \label{eq:kinematic}  
 \frac{D}{D t} \left( r - R_i(z,t) \right) = 0 \quad \text{on } r = R_i(z,t)
-\eeq
+\end{equation}
 and analogously for $R_o(z,t)$.
@@
 Because the BM thickness is much smaller than the vessel length it is possible, on using the lubrication approximation, to derive a 1D model (in $z$) for the BM thickness $h_\textrm{bm}(z,t)$; this takes the form
-\beq
-\frac{\dd}{\dd t} \left( \gamma R_i(z,t) \cdot h_\textrm{bm}(z,t) \right) = \frac{\dd}{\dd z} \left( R_i(z,t) \cdot h_\textrm{bm}(z,t) \cdot K(p_{z}) \cdot p_{z} \right),\label{eq:pvs_model}
-\eeq
-where $\gamma$ is the fluid volume fraction in the BM.
+\begin{equation}
+\begin{split}
+\frac{\dd}{\dd t} \left( \gamma R_i(z,t) \cdot h_\textrm{bm}(z,t) \right) &=\\
+\frac{\dd}{\dd z} \left( R_i(z,t) \right. & \left. \cdot h_\textrm{bm}(z,t) \cdot K(p_{z}) \cdot p_{z} \right),\label{eq:pvs_model}
+\end{split}
+\end{equation}
+where $\gamma$ is the fluid volume fraction in the BM (see also SI~1).
 Equation \eqref{eq:pvs_model} describes the evolution of the BM thickness and can be solved using a finite difference approximation.
-For a detailed derivation see Supplementary Information 1.
+For a detailed derivation see SI~1.
 Another method to arrive at the same equation is derived by considering an infinitesimally small volume of BM $dV = 2\pi \cdot R_i(z,t) \cdot h_\textrm{bm}(z,t) \cdot dz$.
-Then the difference between the flux into and out of the element $dV$ is equal to $\dd (dV)/ \dd t$, which can be written as
-\begin{align}
-2\pi \frac{\dd}{\dd t} \left( \gamma R_i(z,t) \cdot h_\textrm{bm}(z,t) \right) &= \nonumber \\ 
-2\pi \cdot \gamma R_i(z,t) \cdot h_\textrm{bm}(z,t) & \cdot q(z,t) - 2\pi \cdot \gamma R_i(z + dz,t) \cdot h_\textrm{bm}(z + dz,t) \cdot q(z + dz,t),
-\end{align}
-which is equal to the model equation \eqref{eq:pvs_model}.
-The pressure $p(z,t)$ and the displacement of the BM wall $R_i(z,t)$ are, to a very good approximation, determined solely by the arterial pressure (\ie almost independent of the BM flow), see Supplementary Information 2 for more details.
-$R_i(z,t)$ and $p_z$ can be obtained from a numerical artery model described in the next section.
-To model the protein valve that is being proposed here $K(p_z)$ is defined as a step function
+Then the difference between the flux into and out of the element $dV$ is equal to the rate of change of the fluid volume $\dd (\gamma dV)/ \dd t$, which can be written as
+\begin{equation}
+\begin{split}
+\frac{\dd}{\dd t} \left( \gamma R_i(z,t) \cdot h_\textrm{bm}(z,t) \cdot dz \right) &=\\
+& R_i(z,t) \cdot h_\textrm{bm}(z,t) \cdot q_1(z,t) -\\
+R_i(z + dz,t) & \cdot h_\textrm{bm}(z + dz,t) \cdot q_1(z + dz,t),
+\end{split}
+\end{equation}
+which is equal to the model equation \eqref{eq:pvs_model} in the limit $dz \rightarrow 0$.
+The pressure $p(z,t)$ and the displacement of the BM wall $R_i(z,t)$ are, to a very good approximation, determined solely by arterial pressure (\ie almost independent of the BM flow), see SI~2 for more details.
+$R_i(z,t)$ and $p_z$ are obtained from a numerical artery model described in the next section.
+To model the protein valve that is being proposed here, $K(p_z)$ is defined as a step function
 \begin{equation}
@@
 \end{equation}
-chosen such that $K_1 > K_0 > 0$ so that the BM is more permeable to flow of ISF in the negative $z$-direction than in the positive one.
+such that $K_1 > K_0 > 0$ so that the BM is more permeable to ISF flow in the negative $z$-direction than in the positive one.
 
-\section{Numerical Model of a Straight Arterial Section}
+\subsection{Numerical Model of a Straight Arterial Section}
 
-\begin{table}
-\center
-\caption{Geometry parameters of the MCA section modelled using VaMpy \cite{Diem2016a}.\label{tab:geometry}}
-\begin{tabular}{ccc}
-\toprule
+\begin{table}[!t]
+\textbf{\refstepcounter{table}\label{tab:geometry} Table \arabic{table}.}{ Geometry parameters of the MCA section modelled using VaMpy \cite{Diem2016a}. }
+\processtable{ }
+             {\begin{tabular}{ccccl}\toprule
+& MCA & left daughter & right daughter & \\
+&     & vessel        & vessel         & \\
+\midrule
+$R_u$ & \SI{0.14}{\centi\metre} & \SI{0.13}{\centi\metre} & \SI{0.13}{\centi\metre} & upstream radius \\
+$R_d$ & \SI{0.14}{\centi\metre} & \SI{0.08}{\centi\metre} & \SI{0.08}{\centi\metre} & downstream radius \\
+$L$ & \SI{7}{\centi\metre} & \SI{6.5}{\centi\metre} & \SI{6.4}{\centi\metre} & length\\
+\botrule
+\end{tabular}}{}
+\end{table}
+
+\begin{table}[!t]
+\textbf{\refstepcounter{table}\label{tab:simulation} Table \arabic{table}.}{ Simulation parameters for modelling a bifurcation of the MCA using VaMpy \cite{Diem2016a}. }
+\processtable{ }
+             {\begin{tabular}{ccc}\toprule
 Parameter & Value & Physical meaning\\
 \midrule
-\multicolumn{3}{l}{MCA}\\
-\midrule
-$R_u$ & \SI{0.14}{\centi\metre} & upstream radius\\
-$R_d$ & \SI{0.14}{\centi\metre} & downstream radius\\
-$L$ & \SI{7}{\centi\metre} & length\\
-\midrule
-\multicolumn{3}{l}{Left daughter vessel}\\
-\midrule
-$R_u$ & \SI{0.13}{\centi\metre} & upstream radius\\
-$R_d$ & \SI{0.08}{\centi\metre} & downstream radius\\
-$L$ & \SI{6.5}{\centi\metre} & length\\
-\midrule
-\multicolumn{3}{l}{Right daughter vessel}\\
-\midrule
-$R_u$ & \SI{0.13}{\centi\metre} & upstream radius\\
-$R_d$ & \SI{0.08}{\centi\metre} & downstream radius\\
-$L$ & \SI{6.4}{\centi\metre} & length\\
-\bottomrule
-\end{tabular}
+$k_1$ & \SI{2.0e7}{\gram\per\centi\metre\per\square\second} & wall elasticity \\
+$k_2$ & \SI{-22.53}{\per\centi\metre} & parameters, \\
+$k_3$ & \SI{8.65e5}{\gram\per\centi\metre\per\square\second} & see \eqref{eq:ehr} \cite{Olufsen2000,Diem2016a} \\
+$\nu$ & \SI{0.046}{\square\centi\metre\per\second} & kinematic viscosity of blood\\
+$R_1$ & \SI{14130}{\gram\per\centi\metre\tothe{4}\per\second} & First resistance element\\
+$R_2$ & \SI{7200}{\gram\per\centi\metre\tothe{4}\per\second} & Second resistance element\\
+$C_T$ & \SI{2.4752e-3}{\centi\metre\tothe{4}\square\second\per\gram} & Compliance element\\
+$T$ & \SI{0.85}{\second} & cardiac cycle length\\
+$t_c$ & 4 & number of cardiac cycles\\
+$\Delta t$ & \SI{1e-5}{\second} & time step size\\
+$\Delta z$ & \SI{0.1}{\centi\metre} & spatial step size\\
+\botrule
+\end{tabular}}{}
 \end{table}
 
-In order to obtain the input functions $R_i(z,t)$ and $p_z$ for the analytical model developed in the previous section, a section of human MCA was modelled as a 1D axisymmetric tube in a cylindrical coordinate system with cross-sectional lumen area $A(z,t)$ and flux $q(z,t)$ using the Python package VaMpy \cite{Diem2016a}.
-The geometry parameters are listed in \autoref{tab:geometry}.
-Following the derivations in \cite{Olufsen2000,Diem2016} blood flow through arteries is represented by the conservation law
+In order to obtain the input functions $R_i(z,t)$ and $p_z$ for the analytical model developed in the previous section, a section of human MCA was modelled as a 1D axisymmetric tube in a cylindrical coordinate system with cross-sectional lumen area $A(z,t)$ and flux $Q(z,t)$ using the Python package VaMpy \cite{Diem2016a}.
+The geometry parameters are listed in Table~\ref{tab:geometry}.
+Blood flow through arteries and the propagation of arterial pulsations are commonly described by a one-dimensional averaged model \cite{Olufsen2000} (for more details see \cite{Diem2016}).
+It is represented by the conservation law
 \begin{equation}
@@
 \begin{gather*}
-  \boldsymbol{U} = \begin{pmatrix} A(z,t) \\ q(z,t) \end{pmatrix}, \quad \boldsymbol{F} = \begin{pmatrix} q(z,t)\\ \dfrac{q(z,t)^2}{A(z,t)} + f(r_0) \sqrt{A_0(z) A(z,t)} \end{pmatrix}, \quad \boldsymbol{S} = \begin{pmatrix} 0\\ S_1 \end{pmatrix}\\
-  \begin{split}
-    S_1 = -&\frac{2 \pi R(z,t)}{\delta_b \mathrm{Re}} \frac{q(z,t)}{A(z,t)} +\\
-    &\left( 2 \sqrt{A(z,t)} \left( \sqrt{\pi} f(r_0) + \sqrt{A_0(z)} \frac{df(r_0)}{dr_0 } \right) - A(z,t) \frac{df(r_0)}{dr_0} \right) \frac{dr_0(z)}{dz}.
-  \end{split}
+  \boldsymbol{U} = \begin{pmatrix} A(z,t) \\ Q(z,t) \end{pmatrix}, \quad \boldsymbol{F} = \begin{pmatrix} Q(z,t)\\ \dfrac{Q(z,t)^2}{A(z,t)} + f(r_0) \sqrt{A_0(z) A(z,t)} \end{pmatrix}, \quad \boldsymbol{S} = \begin{pmatrix} 0\\ S_1 \end{pmatrix}\\
+S_1 = - \frac{2 \pi R(z,t)}{\delta_b \mathrm{Re}} \frac{Q(z,t)}{A(z,t)} + \left( 2 \sqrt{A(z,t)} \left( \sqrt{\pi} f(r_0) + \sqrt{A_0(z)} \frac{df(r_0)}{dr_0 } \right) - A(z,t) \frac{df(r_0)}{dr_0} \right) \frac{dr_0(z)}{dz}.
 \end{gather*}
-The equation is solved for cross-sectional area $A(z,t)$ of the vessel lumen and flux $q(z,t)$ where $R(z,t) = q(z,t)/A(z,t)$ is the lumen radius, $\delta_b$ is the boundary layer thickness and $\text{Re}$ is the Reynolds number.
-Elasticity of the artery wall is modelled via the quantity $f(r_0) = 4Eh/(3r_0)$, where $E$ is the Young's modulus and $h$ is the wall thickness.
+The equation is solved for cross-sectional area $A(z,t)$ of the vessel lumen and flux through the vessel $Q(z,t)$, where $R(z,t) = (A(z,t)/\pi)^{1/2}$ is the lumen radius, $\delta_b$ is the boundary layer thickness and $\text{Re}$ is the Reynolds number.
+Elasticity of the artery wall is modelled via the quantity $f(r_0) = 4Eh/(3r_0)$, where $E$ is the Young's modulus and $h$ is the wall thickness and $r_0(z)$ is the radius at rest.
 The relationship $Eh/r_0$ is based on compliance estimates
 \begin{equation}
-  \frac{Eh}{r_0} = k_1 \exp \left( k_2 r_0(z) \right) + k_3
+  \frac{Eh}{r_0} = k_1 \exp \left( k_2 r_0(z) \right) + k_3\label{eq:ehr}
 \end{equation}
-with $k_1, k_2, k_3$ constants \cite{Olufsen2000}. Blood pressure $p(z,t)$ is related to $A(z,t)$ and $f(r_0)$ via the state equation
-\beq
+with $k_1, k_2, k_3$, see Table~\ref{tab:simulation} \cite{Olufsen2000}. Blood pressure $p(z,t)$ is related to $A(z,t)$ and $f(r_0)$ via the state equation
+\begin{equation}
   p(z,t) - p_0 = f(r_0) \left( 1 - \frac{A_0(z)}{A(z,t)} \right),
-\eeq
-where $p_0$ is the diastolic pressure and $A_0(z)$ is the lumen radius at rest.
-Blood pressure provides the basis for determining ISF pressure inside the BM via the calculations in Supplementary Information 2.
+\end{equation}
+where $p_0$ is the diastolic pressure and $A_0(z)$ is the lumen radius at rest. Blood pressure provides the basis for determining ISF pressure inside the BM via the calculations in SI~2.
 
 To derive \eqref{eq:conservation} it was necessary to make assumptions about the velocity profile of blood flow through an artery.
-Blood flow is considered pulsatile laminar and vessels are considered slightly tapered, therefore the velocity profile is assumed to be mostly flat with a thin boundary layer $\delta_b = (\nu T/(2\pi))^{0.5}$, such that $\delta_b \ll R(z,t)$ and $\text{Re} \approx 200$ \cite{Olufsen2000}.
-
-\begin{table}
-\center
-\caption{Simulation parameters for modelling a bifurcation of the MCA using VaMpy \cite{Diem2016a}.\label{tab:simulation}}
-\begin{tabular}{ccc}
-\toprule
-Parameter & Value & Physical meaning\\
-\midrule
-$k_1$ & \SI{2.0e7}{\gram\per\centi\metre\per\square\second} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
-$k_2$ & \SI{-22.53}{\per\centi\metre} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
-$k_3$ & \SI{8.65e5}{\gram\per\centi\metre\per\square\second} & wall elasticity parameter (see \cite{Olufsen2000,Diem2016a})\\
-$\nu$ & \SI{0.046}{\square\centi\metre\per\second} & kinematic viscosity of blood\\
-$R_1$ & \SI{14130}{\gram\per\centi\metre\tothe{4}\per\second} & First resistance element\\
-$R_2$ & \SI{7200}{\gram\per\centi\metre\tothe{4}\per\second} & Second resistance element\\
-$C_T$ & \SI{2.4752e-3}{\centi\metre\tothe{4}\square\second\per\gram} & Compliance element\\
-$T$ & \SI{0.85}{\second} & cardiac cycle length\\
-$t_c$ & 4 & number of cardiac cycles simulated\\
-$\Delta t$ & \SI{1e-5}{\second} & time step size\\
-$\Delta z$ & \SI{0.1}{\centi\metre} & spatial step size\\
-\bottomrule
-\end{tabular}
-\end{table}
+Blood flow is considered pulsatile laminar and vessels are considered slightly tapered, therefore the velocity profile is assumed to be mostly flat with a thin boundary layer $\delta_b = (\nu T/(2\pi))^{0.5}$, such that $\delta_b \ll R(z,t)$ \cite{Olufsen2000}.
 
 This system of equations can be solved numerically using Richtmyer's two-step version of the Lax-Wendroff method, as done by others \cite{Olufsen2000,LeVeque1992,Kolachalama2007}.
-We use the code from \cite{Diem2016a}, which is an implementation of the Lax-Wendroff method in Python, to implement a model of the MCA and its first bifurcation.
-All simulation parameters are listed in \autoref{tab:simulation}.
-The MCA radius was taken from \cite{Cousins2014} and following their estimate that generally the length of an artery is $50 \cdot R_u$ lengths were calculated accordingly.
-The radii of the daughter vessels were estimated using a scaling factor of 0.91, also obtained from \cite{Cousins2014}.
-The length of the right daughter vessel was shortened slightly to avoid reflective waves cancelling one another out, as also done in \cite{Kolachalama2007}.
+We use the Python module from \cite{Diem2016a} to implement a model of the MCA and its first bifurcation.
+All simulation parameters are listed in Table~\ref{tab:simulation}.
+The MCA geometry data was taken from \cite{Cousins2014}, estimating lengths by $50 \cdot R_u$.
+The radii of the daughter vessels were estimated using a scaling factor of 0.91 \cite{Cousins2014}.
+The length of the right daughter vessel was shortened slightly to avoid reflective waves cancelling one another out \cite{Kolachalama2007}.
 A three-element Windkessel (3WK) model is used at the outlet whose parameters were obtained from \cite{Olufsen2002}.
 
-\begin{figure}
-\centerline{\includegraphics{figures/inlet.png}}
-\caption{Inlet function for MCA simulations obtained from patient-specific measurements used to represent one period of \SI{0.85}{\second} of the pulse wave.
-The final ten peaks of the velocity time series were averaged in the Fourier space and adjusted such that its range of values aligns with those reported in \cite{Olufsen2002}.
-The resulting inlet function is smooth and serves as the inlet boundary condition for the MCA simulations using VaMpy \cite{Diem2016a}.\label{fig:inlet}}
-\end{figure}
+At the inlet flux is prescribed using patient-specific flow velocity measurements collected from the MCA of a healthy adult male using Doppler Sonography ultrasound.
+The velocity data was converted to volumetric flux values within the range reported in \cite{Olufsen2002}.
+To obtain a smooth inlet function the final ten peaks of the time series were averaged in the Fourier space.
+The resulting inlet boundary condition is shown in Figure~\ref{fig:inlet}.
 
-At the inlet the flux is prescribed directly.
-Patient-specific flow velocity measurements were collected from the MCA of a healthy adult male using Doppler Sonography ultrasound.
-The velocity data was adjusted to obtain volumetric flux values within the range reported in \cite{Olufsen2002}.
-To obtain a smooth inlet function the final ten peaks of the time series were averaged in the Fourier space.
-The resulting inlet boundary condition is shown in \autoref{fig:inlet}.
+\subsection{Perivascular Drainage through the MCA}
 
-\begin{figure}
-\centerline{\includegraphics{figures/mca.png}}
-\caption{Flux (a) and pressure (b) in the MCA.
-Simulations were performed using the VaMpy Python package \cite{Diem2016a} and the simulation parameters are listed in Table~\ref{tab:simulation}.
-Wall stiffness is high in small blood vessels, therefore pressure gradients in time are steep.\label{fig:mca}}
-\end{figure}
-
-MCA simulations were run for four cardiac cycles to allow the system to settle.
-\autoref{fig:mca} shows flow and pressure in the MCA.
+In the previous sections the models governing blood flow through the MCA and ISF flow through the artery wall have been introduced.
+Here, the results of the previous sections are coupled and used to calculate ISF flow through the BM with and without a valve mechanism to show its necessity to achieve net reverse drainage.
+At the same time it is shown that blood pressure driven perivascular drainage is far too slow to provide any meaningful flow, thereby disproving the popular hypothesis that arterial pulsations provide the major driving force for perivascular drainage.
+MCA simulations were run for eight cardiac cycles to allow the system to settle.
+An averate human heartbeat lasts \SI{0.85}{\second} and the total simulation time was \SI{6.8}{\second} (see Table~\ref{tab:simulation}).
+In the results presented here only the final heartbeat is shown.
+Figure~\ref{fig:mca} shows blood pressure (a), wall displacement (b) and ISF pressure (c) in the MCA.
 Because the radius of the MCA is very small its wall stiffness is high.
@@
 
-\section{Perivascular Drainage through the MCA}
+Figure~\ref{fig:mca}b shows the displacement of the artery wall due to the pressure pulse.
+To calculate the width of the BM using the model equation \eqref{eq:pvs_model}, the input function $R_i(z,t)$ is the sum of the radius at rest and the wall displacement, as depicted in Figure~\ref{fig:pvs_model}.
+Figure~\ref{fig:h}a shows the width of the BM $h_\textrm{bm}(z,t)$ as calculated from \eqref{eq:pvs_model} over the length of the artery at a number of time points, while Figure~\ref{fig:h}b shows $\partial h_\textrm{bm}(z,t)/ \partial z$ for the same time points.
+The BM width only varies minimally around its initial width and thus flow rates of ISF are very small.
 
-In the previous sections the models governing blood flow through the MCA and ISF flow through the artery wall have been introduced.
-Here, the results of the previous sections are coupled and used to calculate ISF flow through the BM with and without a valve mechanism to show its necessity to achieve net reverse drainage.
-At the same time it is shown that blood pressure driven perivascular drainage is far too slow to provide any meaningful flow, thereby disproving the popular hypothesis that arterial pulsations provide the major driving force for perivascular drainage.
-The numerical model of the MCA was run over four heartbeats to ensure convergence of the results.
-An averate human heartbeat lasts \SI{0.85}{\second} and the total simulation time was \SI{3.4}{\second} (see Table~\ref{tab:simulation}).
-In the results presented here only the final heartbeat is shown.
+The results show that a valve mechanism is required to drive perivascular drainage in the reverse direction of the blood flow with an average flux of \SI{-1.12e-3}{\cubic\micro\metre\per\second} for a ratio of $K_0/K_1 = 0.01$, while it is \SI{6.33e-2}{\cubic\micro\metre\per\second} for $K_0/K_1 = 1.0$.
+While the results confirm the necessity of a strong valve mechanism under pulse driven flow they also indicate that arterial pulsations have too long a wavelength to drive significant perivascular drainage.
+Drainage velocity is fastest without a valve mechanism (\SI{1.81e-2}{\micro\metre\per\second}, and then it is in the wrong direction), but even in that case it is almost three orders of magnitude slower than the expected value of \SI{8.33}{\micro\metre\per\second} \cite{Carare2008}.
+The valve mechanism only guarantees net reverse drainage for a permeability ratio $K_0/K_1 < \SI{2.74e-2}{}$, where $K_1$ is chosen such that $k = \SI{1e-10}{\square\centi\metre}$ is consistent with values for other tissues in the body and $\mu = \SI{1.5e-3}{\pascal\second}$ \cite{Heppell2013}.
+Figure~\ref{fig:valve_test} illustrates flux through the BM for a number of ratios $K_0/K_1$ and for different positions of the BM $r = r_0 + \eta h$.
+Reverse flow is achieved for $K_0/K_1 < \SI{2.74e-2}{}$.
+Drainage flux values are fastest when the BM is placed close to the arterial lumen ($\eta = 0$) due to the elasticity of the artery wall.
 
-\begin{figure}
-\centerline{\includegraphics{figures/Ri.png}}
-\caption{Displacement of the artery wall due to the pressure pulse in the MCA.
-The displacement is calculated from the linear elasticity approximation derived in Supplementary Information 2.
-The input function $R_i(z,t)$ to the BM model equation \eqref{eq:pvs_model} is obtained by adding the radius at rest $a$ (see geometry depiction in Figure~\ref{fig:pvs_model}).\label{fig:ri}}
-\end{figure}
+The minute flux values are a result of the very small variations in BM width (see Figure~\ref{fig:h}) over the length of the MCA, in fact they are less than \SI{1}{\angstrom}.
+This value is far too small to be realistic and demonstrates that arterial pulsations cannot be strong enough to drive perivascular drainage through the BM.
+To illustrate it in another way, the total ISF volume in a human brain is \SI{280}{\milli\litre} \cite{Weller2015a}.
+At a drainage rate of \SI{-1.12e-3}{\cubic\micro\metre\per\second} per cerebral artery the total turnover time for ISF would be \SI{4.83e11}{} days, much more than a human life span.
 
-\begin{figure}
-\centerline{\includegraphics{figures/h_dh.png}}
-\caption{BM width $h_\textrm{bm}(z,t)$ at several time points during the cardiac cycle (a) and its gradient along the $z$-axis at the same time points (b).
-Pressure pulse driven variations of the BM width are small compared to the initial width of \SI{200}{\nano\metre} and thus flow through the BM is slow.\label{fig:h}}
-\end{figure}
+\section{Experimental assessment of intramural perivascular drainage of \Ab}
 
-Figure~\ref{fig:ri} shows the displacement of the artery wall due to the pressure pulse.
-To caluculate the width of the BM using the model equation \eqref{eq:pvs_model}, the input function $R_i(z,t)$ is the sum of the radius at rest and the wall displacement, as depicted in Figure~\ref{fig:pvs_model}.
-Figure~\ref{fig:h}a shows the width of the BM $h_\textrm{bm}(z,t)$ as calculated from \eqref{eq:pvs_model} over the length of the artery at a number of time points, while Figure~\ref{fig:h}b shows $\partial h_\textrm{bm}(z,t)/ \partial z$ for the same time points.
-The BM width only varies by approximately $\pm \SI{0.01}{\nano\metre}$ around its initial width and thus results in very low flow rates of ISF.
+In addition to performing simulations to assess the feasibility of arterial pulsations providing the main driving mechanism for IPAD, mouse model experiments were performed to confirm the results \textit{in vivo}.
+Experiments on mice were carried out on ten week old wild-type C57B16 mice ($n = 4$).
+All mice were kept on a standard \SI{12}{\hour} light/dark cycle and allowed food and water \textit{ad libitum}.
+All experiments were carried out in accordance with animal care guidelines stipulated by the Animal Care and Use Committee at the University of Southampton and the Home Office (PPL 20/2095).
 
-\begin{figure}
-\centerline{\includegraphics{figures/bm_pressure.png}}
-\caption{ISF pressure inside the BM of the MCA.
-To obtain ISF pressure the radial stresses inside the wall were calculated from the blood pressure inside the artery.
-ISF pressure is directly proportional to the radial stress in the wall at the location of the BM (see Supplementary Information 2).
-Flow velocity was then calculated using Darcy's law.\label{fig:bm_pressure}}
-\end{figure}
+To minimise the effects of anaesthesia on mouse cardiovascular function, isoflurane was chosen as the anaesthetizing agent.
+To reduce arterial pulsations, the beta-blocker Atenolol (Ate, \SI{10}{\milli\gram\per\kilo\gram}) was administered intraperitoneally (IP).
+Mouse vital parameters were monitored throughout the procedure using a MouseOX Plus and an infrared mouse thigh sensor (STARR Life Science, Holliston, MA, USA).
+Thermal regulation was maintained using a heat pad and rectal probe (BASi, West Lafayette, IN, USA).
+All intracerebral injections were performed using a \SIrange{1}{5}{\micro\litre} Hirschmann microcapillary pipette (Sigma UK) with a tip adjusted to a diameter of $<$ \SI{50}{\micro\metre} using a Sutter P97 Flaming Brown Pipette puller.
 
-\begin{figure}
-\centerline{\includegraphics{figures/valve_test.png}}
-\caption{ISF velocity at $r = r_0 + xh$.
-Drainage occurs in the reverse direction of the blood flow for $K_0/K_1 = \SI{1.6e-2}{}$ and is fastest at $r = r_0 + 0.1h$.
-The fastest net reverse drainage is \SI{-7.35e-6}{\micro\metre\per\second}, which is six orders of magnitude smaller than the expected value of \SI{-8.33}{\micro\metre\per\second}.\label{fig:valve_test}}
-\end{figure}
+Mice were anaesthesized with isoflurane, placed prone on a stereotaxic instrument and secured with head adaptors.
+Baseline measurements of oxygen saturation, heart rate, breath rate and pulse distension were taken before administering Ate.
+A volume of \SI{0.5}{\micro\litre} of \Ab$_{1-40}$ HiLyte Fluor 555 (Cambridge Bioscience, UK) was stereotaxically injected into the left hippocampus at a rate of \SI{0.2}{\micro\litre\per\minute}.
+Injection pipettes were left \textit{in situ} for \SI{2}{\minute} to prevent reflux.
+Mice were euthanized \SI{5}{\minute} after withdrawal of the injection pipette through overdose with pento-barbital (\SI{200}{\milli\gram\per\kilo\gram}).
+Mice were then intracardially perfused with \SI{0.01}{\Molar} phosphate buffered saline (PBS), pH 7.2 followed by \SI{4}{\percent} paraformaldehyde in \SI{0.01}{\Molar} PBS, pH 7.2.
+Brains were removed, post fixed overnight in fresh fixative and then processed for immunohistochemistry and image analysis.
+For controls, ten week old wild-type C57B16 mice were processed following the same method, but without the administration of Ate ($n = 3$).
+Detailed materials and methods can be found in SI~3.
 
-ISF pressure inside the BM (at $r = r_0 + 0.5h$) is shown in \autoref{fig:bm_pressure}.
-The results show that a valve mechanism is required to drive perivascular drainage in the reverse direction of the blood flow with an average flow velocity of \SI{-3.54e-6}{\micro\metre\per\second} for a ratio of $K_0/K_1 = 0.01$, while it is \SI{7.91e-4}{\micro\metre\per\second}.
-While the results confirm the necessity of a strong valve mechanism under pulse driven flow they also indicate that arterial pulsations are not strong enough to drive perivascular drainage.
-Drainage velocity is fastest without a valve mechanism (and then it is in the wrong direction), but even in that case it is four orders of magnitude slower than the expected value of \SI{8.33}{\micro\metre\per\second} \cite{Carare2008}.
-The valve mechanism only guarantees net reverse drainage for a permeability ratio $K_0/K_1 < \SI{1.6e-2}{}$, where $K_1$ is chosen such that $k = \SI{1e-10}{\square\centi\metre}$ is consistent with values for other tissues in the body and $\mu = \SI{1.5e-3}{\pascal\second}$ \cite{Heppell2013}.
+\subsection{Immunohistochemistry}
+
+Brains were cut into \SI{20}{\micro\metre} coronal sections using a freezing microtome (Leica) and the injection site located by examining on a Zeiss Axioskop 2 fitted with a rhodamine filter.
+Sections of \SI{400}{\micro\metre} and \SI{800}{\micro\metre} thickness both posterior and anterior to the injection site were selected for immunohistochemistry.
+Sections were blocked in \SI{15}{\percent} goat serum (Sigma 9023) for \SI{1}{\hour} at room temperature (RT).
+They were then incubated in primary antibodies diluted in \SI{0.01}{\Molar} phosphate buffered saline \SI{0.1}{\percent} triton x100 (PBSt).
+Primary antibodies comprised rabbit anti-collagen IV (AbCam, ab6586) 1/400 and mouse anti-smooth muscle actin (SMA) FITC conjugated (Sigma F3777) 1/200.
+Sections were incubated overnight at \SI{4}{\celsius} in a moist chamber.
+Following washing 3x \SI{10}{\minute} \SI{0.01}{\Molar} PBS sections were incubated in goat anti-rabbit Alexa Fluor 633 (ThermoFisher Scientific A-21070) 1/200 PBSt.
+
+\subsection{Image Analysis}
+
+Images of the whole hippocampus were captured on a Leica SP8 confocal microscope fitted with LASX software in tile scanning mode with sequential imaging.
+Once captured, images were exported as TIFF files and uploaded into Adobe Photoshop CS6.
+An area of hippocampus was drawn extending transversely from CA3 pyramidal cells to the apex of the blades of the granule cell layer and from the ventricle to \textit{stratum oriens} in the superior/inferior axis.
+\Ab containing vessels were defined according to size and presence of SMA: capillaries had a diameter of $\leq$ \SI{10}{\micro\metre}, veins had no SMA and a diameter of $>$ \SI{10}{\micro\metre}.
+The number of vessels was expressed per \SI{}{\square\milli\metre} of hippocampus.
+Statistical analysis was performed by an independent t-test with a confidence interval of \SI{95}{\percent}.
+
+\subsection{Results}
+
+Within \SI{5}{\minute} of injections into the hippocampus, fluorescent \Ab was observed diffusely in the hippocampus and co-localising with collagen IV with collagen IV in the walls of capillaries and arteries, with very few veins involved in both Ate-treated and control mice (Figure~\ref{fig:confocal}).
+Application of an independent t-test to the data revealed no significant differences between the saline and Ate treated mice either anterior or posterior to the injection site (Figure~\ref{fig:graph}).
 
@@
 The results show that the valve mechanism as implemented in our analytical model of the BM guarantees net reverse flow inside the BM under physiological blood flow conditions.
-The valve mechanism is required to be strong, \ie the ratio of $K_0/K_1$ needs to be small.
-Additionally, the model shows that a valve mechanism is neccessary for net reverse flow to occur within the BM, confirming results from previous studies that did not include a valve mechanism and concluded the existence of some attachment mechanism \cite{Schley2006,Wang2011}.
-Net drainage additionally depends on the distance of the BM from the artery lumen.
+The valve mechanism is required to be strong, \ie with a small ratio $K_0/K_1$, and necessary for net reverse flow to occur within the BM.
+This result agrees with previous modelling approaches that did not achieve net reverse drainage \cite{Schley2006,Wang2011}.
+The velocity of drainage additionally depends on the distance of the BM from the artery lumen.
 The closer the BM is placed to the lumen the faster net reverse drainage is, but variations remain within the same order of magnitude.
 This is due to the relatively large thickness of cerebral artery walls and their elastic properties.
-The hypothesis that the BM utilises a unidirectional valve mechanism is plausible as the general lymphatic system of the body also has one and its existence has previously been speculated \cite{Weller2010,Schley2006,Heppell2013}.
+The hypothesis that the BM utilises a unidirectional valve mechanism is plausible as the general lymphatic system of the body also has one and its existence has previously been discussed \cite{Weller2010,Schley2006,Heppell2013}.
 
@@
 
-The results from this study showed that the average drainage velocity of ISF is \SI{-3.54e-6}{\micro\metre\per\second} at $r = r_0 + 0.5h$ and $K_0/K_1 = 0.01$.
-This value indicates a net reverse flow within the BM, although it is six orders of magnitude smaller than what would be expected \cite{Carare2008}.
+The results from this study showed that the average drainage flux of ISF is \SI{-1.12e-2}{\cubic\micro\metre\per\second} at $r = r_0 + 0h$ and $K_0/K_1 = 0.01$.
+This value indicates a net reverse flow within the BM, although it is eight orders of magnitude smaller than the flux of \SI{0.001}{\cubic\milli\metre\per\second} estimated by \cite{Arbel-Ornath2013}.
 The results therefore suggest that arterial pulsations are not powerful enough to drive perivascular drainage of ISF through the BM.
-This is most likely due to the very long wavelength of the arterial pulsations compared to the artery section considered, which results in very small pressure gradients along the vessel length (\autoref{fig:mca}b).
-It appears that other forces are neccessary to produce the pressure gradients required to push fluid through the BM and an interesting candidate for such forces is vasomotion of smooth muscle cells \cite{DiMarco2015}.
+This is due to the very long wavelength of the arterial pulsations compared to the artery section considered, which results in very small pressure gradients along the vessel length (Figure~\ref{fig:mca}b).
+These results are confirmed here by analysing the pattern of IPAD in mice after administration of pulse modulators.
+No significant differences were found between mice that received the pulse modulator and control mice.
+It appears that other forces are neccessary to produce the pressure gradients required to push fluid through the BM and an interesting candidate for such forces is the contraction of vascular smooth muscle cells, as noted in \cite{DiMarco2015} and modelled in detail by \cite{Aldea2016}.
 
@@
 Other studies, which have considered arterial pulsations as driving mechanisms have either concluded that ISF flow in the reverse direction to the blood flow requires some form of attachment mechanism \cite{Schley2006,Wang2011} or require very specific protein movements giving rise to valve-like behaviour \cite{Sharp2015}.
-Our model is the first to suggest a very general valve mechanism that reliably produces net reverse drainage along the BM.
-While the valve model of the BM provides a phyisologically feasible mechanism to ensure net reverse drainage of ISF it is insufficient to drive the physiologically observed flow (by several orders of magnitude).
-This result is therefore significant as it disproves the widely accepted arterial pulsation hypothesis of perivascular drainage of \Ab from the brain \cite{Weller2009,Carare2008,Hawkes2011,Morris2014,Schley2006,Attems2011,Wang2011,Iliff2012,Asgari2015,Sharp2015,Weller2015a} and suggests that other forces have to be considered instead.
+Our model suggests a very general valve mechanism that reliably produces net reverse drainage along the BM.
+While the valve model of the BM provides a phyisologically feasible mechanism to ensure net reverse drainage of ISF, it is too small by several orders of magnitude to drive the physiologically observed IPAD.
+This mathematical model as well as experimental study disproves the widely accepted arterial pulsation hypothesis of perivascular drainage of \Ab from the brain \cite{Weller2009,Carare2008,Hawkes2011,Morris2014,Schley2006,Attems2011,Wang2011,Iliff2012,Asgari2015,Sharp2015,Weller2015a} and suggests that other forces must be considered for efficient IPAD and for therapeutic strategies in AD.
 
-% ----------------------------------------------------------------------------
-% Acknowledgements
-% ----------------------------------------------------------------------------
-\section*{Acknowledgements}
+\section*{Disclosure/Conflict-of-Interest Statement}
+%Frontiers follows the recommendations by the International Committee of Medical Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require that all financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed.
+If no such relationship exists, authors will be asked to declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
+When disclosing the potential conflict of interest, the authors need to address the following points:
+%•	Did you or your institution at any time receive payment or services from a third party for any aspect of the submitted work?
+%•	Please declare financial relationships with entities that could be perceived to influence, or that give the appearance of potentially influencing, what you wrote in the submitted work.
+%•	Please declare patents and copyrights, whether pending, issued, licensed and/or receiving royalties relevant to the work.
+%•	Please state other relationships or activities that readers could perceive to have influenced, or that give the appearance of potentially influencing, what you wrote in the submitted work.
 
-This work was supported by an EPSRC Doctoral Training Centre grant (EP/G03690X/1) and EPSRC Doctoral Training Partnership grant (EP/N509747/1).
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
+
+\section*{Author Contributions}
+%When determining authorship the following criteria should be observed:
+%•	Substantial contributions to the conception or design of the work; or the acquisition, analysis, or interpretation of data for the work; AND
+%•	Drafting the work or revising it critically for important intellectual content; AND
+%•	Final approval of the version to be published ; AND
+%•	Agreement to be accountable for all aspects of the work in ensuring that questions related to the accuracy or integrity of any part of the work are appropriately investigated and resolved.
+%Contributors who meet fewer than all 4 of the above criteria for authorship should not be listed as authors, but they should be acknowledged. (http://www.icmje.org/roles_a.html)
+
+A.K.D.
+and G.R.
+designed and analysed BM model and elasticity analysis, A.K.D., N.W.B and G.R.
+designed and analysed numerical models, R.O.C and M.M.S.
+designed and performed mouse experiments, A.K.D., N.W.B, M.M.S., R.O.C.
+and G.R.
+wrote the paper.
+
+\section*{Acknowledgments}
 The authors thank Tony Birch for carrying out the middle cerebral artery flow velocity measurements, Roxana Aldea for our discussions on the appropriate equations for artery wall elasticity and Roy O.
@@
 
-% ----------------------------------------------------------------------------
-% Appendix
-% ----------------------------------------------------------------------------
-%\section*{Appendix}
-%\appendix
-%\input{mathanalysis}
-%\input{airystress}
+\textit{Funding\textcolon} This work was supported by an EPSRC Doctoral Training Centre grant (EP/G03690X/1) and EPSRC Doctoral Training Partnership grant (EP/N509747/1).
 
-% ----------------------------------------------------------------------------
-% References
-% ----------------------------------------------------------------------------
-\addcontentsline{toc}{section}{\bibname}
-
+\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
+%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
 \begin{thebibliography}{10}
@@
 Weller~RO, Carare~RO and Boche~D (2009) Amyloid: Vascular and Parenchymal. {\em Encyclpedia of Neuroscience}, ed Squire~LR (Oxford Academic Press), pp.~355--362.
-
-\bibitem{Weller2009a}
-Weller~RO, Preston~SD, Subash~M and Carare~RO (2009) Cerebral amyloid angiopathy in the aetiology and immunotherapy of Alzheimer's disease. {\em Alzheimer's~Res.~Ther.} 1(2):6.
 
@@
 \bibitem{Heppell2013}
-Heppell~C, Richarson~G and Roose~T (2013) A Model for Fluid Drainage by the Lymphatic System. {\em Bull.~Math.~Biol.} 75(1):49--81.
+Heppell~C, Richardson~G and Roose~T (2013) A model for interstitial drainage through a sliding lymphatic valve. {\em Bull.~Math.~Biol.} 77(6):1101--1131.
 
@@
 Di~Marco~LY, Farkas~E, Martin~C, Venneri~A and Frangi~AF (2015) Is Vasomotion in Cerebral Arteries Impaired in Alzheimer's Disease? {\em J.~Alzheimers~Dis.} 46(1):35--53.
+
+\bibitem{Aldea2016}
+Aldea~RA, Carare~RO, Weller~RO, Richardson~G ({\em in preparation}) Vascular smooth muscle cells as the motive force for perivascular drainage of solutes from the brain. 
 \end{thebibliography}
 
+%%% Upload the *bib file along with the *tex file and PDF on submission if the bibliography is not in the main *tex file
+
+\section*{Figures}
+
+%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
+%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.jpg and logo2.eps) are in the Frontiers LaTeX folder
+%%% If using *.tif files convert them to .jpg or .png
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=10cm]{figures/pvs_model.pdf}
+\textbf{\refstepcounter{figure}\label{fig:pvs_model} Figure \arabic{figure}.}{Geometry of the perivascular drainage pathways model.
+The length $L$ of a typical artery in the brain can be assumed to be much greater than the width $W$ of the perivascular drainage pathways.
+Therefore we can approximate the flow as a lubrication model.
+The pulsation of the capillary is described by the function $h(z,t)$.}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=10cm]{figures/inlet.png}
+\textbf{\refstepcounter{figure}\label{fig:inlet} Figure \arabic{figure}.}{Inlet function for MCA simulations obtained from patient-specific measurements used to represent one period of \SI{0.85}{\second} of the pulse wave.
+The final ten peaks of the velocity time series were averaged in the Fourier space and adjusted such that its range of values aligns with those reported in \cite{Olufsen2002}.
+The resulting inlet function is smooth and serves as the inlet boundary condition for the MCA simulations using VaMpy \cite{Diem2016a}.}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=10cm]{figures/mca.pdf}
+\textbf{\refstepcounter{figure}\label{fig:mca} Figure \arabic{figure}.}{Blood pressure (a), wall displacement (b) and the resulting ISF pressure (c) in the MCA.
+Simulations were performed using the VaMpy Python package \cite{Diem2016a} and the simulation parameters are listed in Table~\ref{tab:simulation}.
+Wall stiffness is high in small blood vessels, therefore pressure gradients in time are steep.
+Wall displacement was calculated from the linear elasticity approximation derived in SI~2.
+The input function $R_i(z,t)$ to the BM model equation \eqref{eq:pvs_model} is obtained by adding the radius at rest $a$ (see geometry depiction in Figure~\ref{fig:pvs_model}).}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=10cm]{figures/h_dh.png}
+\textbf{\refstepcounter{figure}\label{fig:h} Figure \arabic{figure}.}{BM width $h_\textrm{bm}(z,t)$ at several time points during the cardiac cycle (a) and its gradient along the $z$-axis at the same time points (b).
+Pressure pulse driven variations of the BM width are small compared to the initial width of \SI{200}{\nano\metre} and thus flow through the BM is slow.}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=10cm]{figures/valve_test.png}
+\textbf{\refstepcounter{figure}\label{fig:valve_test} Figure \arabic{figure}.}{ISF flux (positive values in the direction of arterial flow) averaged over one cycle as a function of BM position $\eta$ in the arterial wall.
+Here the BM is at $r = r_0 + \eta h$ such that $\eta = 0$ represents a BM immediately adjacent to the lumen and $\eta = 1$ represents a BM on the outer wall of the artery.
+Note that reverse flow is only achieved by ratios of $K_0/K_1 < \SI{2.74e-2}{}$.
+The fastest net reverse drainage shown here (for $K_0/K_1 = \SI{2.68e-2}{}$ and on the arterial lumen $\eta = 0$) is \SI{-2.64e-5}{\cubic\micro\metre\per\second}.}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=10cm]{figures/confocal.jpg}
+\textbf{\refstepcounter{figure}\label{fig:confocal} Figure \arabic{figure}.}{Composite tile scan of confocal images of the distribution of intracerebrally injected \Ab, in relation to collagen IV and SMA.
+Immunostaining \Ab: red, SMA: green, collagen IV: blue.
+Arrows are pointing to arteries with \Ab in the intramural perivascular drainage pathways.}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=10cm]{figures/graph.png}
+\textbf{\refstepcounter{figure}\label{fig:graph} Figure \arabic{figure}.}{Graphs showing the number of arteries, veins and capillaries with \Ab in their walls in atenolol treated mice compared to control mice.
+SEM: standard error of mean.}
+\end{figure}
+
+%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.
+
+%%% Frontiers will add the figures at the end of the provisional pdf automatically %%%
+
+%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.
+
 \end{document}

### 2017-07-18 si01.tex
frontiers submission

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when producing your supplementary material for your Frontiers article.                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 2.1 Generated 2015/05/22 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass{frontiers_suppmat} % for all articles

\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}
\usepackage{siunitx}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqn}{\begin{equation*}}
\newcommand{\eeqn}{\end{equation*}}
\newcommand{\beqs}{\begin{gather}}
\newcommand{\eeqs}{\end{gather}}
\newcommand{\dd}{\partial}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\vareps}{\varepsilon}
\newcommand{\q}{\boldsymbol{q}}
\newcommand{\vel}{\boldsymbol{u}}
\newcommand{\disp}{\boldsymbol{v}}
\newcommand{\norm}{\boldsymbol{n}}
\newcommand{\ra}{\rightarrow}
\newcommand{\hatz}{\hat{z}}
\newcommand{\hatt}{\hat{t}}
\newcommand{\taurr}{\tau_{rr}}
\newcommand{\taurt}{\tau_{r\theta}}
\newcommand{\tautt}{\tau_{\theta\theta}}
\newcommand{\tauzz}{\tau_{zz}}
\newcommand{\taurz}{\tau_{rz}}
\newcommand{\taurrl}{\tau_{rr,0}}
\newcommand{\taurtl}{\tau_{r\theta,0}}
\newcommand{\tauttl}{\tau_{\theta\theta,0}}
\newcommand{\tauzzl}{\tau_{zz,0}}
\newcommand{\taurzl}{\tau_{rz,0}}
\newcommand{\taurre}{\tau_{rr,1}}
\newcommand{\taurte}{\tau_{r\theta,1}}
\newcommand{\tautte}{\tau_{\theta\theta,1}}
\newcommand{\tauzze}{\tau_{zz,1}}
\newcommand{\taurze}{\tau_{rz,1}}
\newcommand{\err}{e_{rr}}
\newcommand{\ert}{e_{r\theta}}
\newcommand{\ett}{e_{\theta\theta}}
\newcommand{\ezz}{e_{zz}}
\newcommand{\erz}{e_{rz}}
\newcommand{\D}{\Delta}
\newcommand{\Rey}{\mathrm{Re}\xspace}
\newcommand{\Airy}{\mathcal{A}}
\newcommand{\wrt}{w.~r.~t.}
\newcommand{\ie}{i.~e.}

% Leave a blank line between paragraphs in stead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Diem {et~al.}} %use et al only if is more than 1 author
\def\Authors{Alexandra K. Diem, Neil W. Bressloff, Roxana O. Carare, Matthew MacGregor Sharp and Giles Richardson}
% The Corresponding Author should be marked with an asterisk
% Provide email of the corresponding author
\def\corrAuthor{A.~K.~Diem}
\def\corrEmail{A.K.Diem@soton.ac.uk}

\begin{document}
\onecolumn
\firstpage{1}

\title[Supplementary Material]{{\helveticaitalic{Supplementary Material}}:
\\ \helvetica{Arterial pulsations cannot drive intramural periarterial drainage: Significance for Alzheimer's disease}} %Please insert the title of your article here

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{Corresponding Author2: email2@uni2.edu}

\maketitle

\section{Mathematical Analysis of flow through the basement membrane}

This section contains the mathematical derivation of the BM valve model equation. The derivation provided here follows are more analytical approach, while the main text contains a more intuitive derivation.

Recall from the main text the dimensionless governing equations and boundary conditions
\begin{gather}
u_z(r,z,t) = -K(p_z) \frac{\dd p(r,z,t)}{\dd z}\label{eq:uz}\\
u_r(r,z,t) = - \frac{1}{\vareps_L} K(p_z) \frac{\dd p(r,z,t)}{\dd r}\\
\frac{\dd u_z(r,z,t)}{\dd z} + \frac{1}{r} \frac{\dd}{\dd r} \left( ru_r(r,z,t) \right) = 0\label{eq:app_continuity}\\
\frac{\dd R_i(z,t)}{\dd t} = u_r(r,z,t) - u_z(r,z,t) \frac{\dd R_i(z,t)}{\dd z} \qquad \text{on } r = R_i(z,t)\label{eq:ri}\\
\begin{split}
\frac{\dd R_i(z,t)}{\dd t} + \vareps_R \frac{\dd h(z,t)}{\dd t} &=\\
u_r(r,z,t) -& u_z(r,z,t) \left( \frac{\dd R_i(z,t)}{\dd z} + \vareps_R \frac{\dd h(z,t)}{\dd t} \right) \qquad \text{on } r = R_o(z,t).\label{eq:ro}
\end{split}
\end{gather}
Note that here the equations are expressed in terms of velocity $\boldsymbol{u}$ instead of flux $\boldsymbol{q}$. Solving Equation~\ref{eq:app_continuity} for the given boundary condition provides values for the pressure gradient $p_z = \dd p/\dd z$ inside the BM, which is then used to calculate ISF flux through the BM. Performing the variable transformation $(z, r, t) \ra (\hatz, \rho, \hatt)$, with $0 \leq \rho \leq h(z,t)$, such that
\begin{gather*}
z = \hatz\\
r = R_i(z,t) + \vareps_R \rho\\
t = \hatt
\end{gather*}
and thus
\begin{gather*}
\frac{\dd}{\dd z} = \frac{\dd}{\dd \hatz} - \frac{1}{\vareps_R} \frac{\dd R_i(z,t)}{\dd \hatz} \frac{\dd}{\dd \rho}\\
\frac{\dd}{\dd r} = \frac{1}{\vareps_R} \frac{\dd}{\dd \rho}\\
\frac{\dd}{\dd t} = \frac{\dd}{\dd \hatt} - \frac{1}{\vareps_R} \frac{\dd R_i(z,t)}{\dd \hatt} \frac{\dd}{\dd \rho}.
\end{gather*}
Application of the variable transformation to the governing equations (Equation~\ref{eq:uz} to Equation~\ref{eq:app_continuity}) results in
\begin{gather}
\hat{u}_{\hatz}(\hatz,\rho,\hatt) = -K(p_{\hatz} - \frac{1}{\vareps_R} \frac{\dd R_i(z,t)}{\dd \hatz} p_{\rho}) \left( \frac{\dd p(\hatz,\rho,\hatt)}{\dd \hatz} - \frac{1}{\vareps_R} \frac{\dd R_i(z,t)}{\dd \hatz} \frac{\dd p(\hatz,\rho,\hatt)}{\dd \rho} \right)\\
\hat{u}_{\rho}(\hatz,\rho,\hatt) = - \frac{1}{\vareps_R \vareps_L^2} K(p_{\hatz} - \frac{1}{\vareps_R} \frac{\dd R_i(z,t)}{\dd \hatz} p_{\rho}) \frac{\dd p(\hatz,\rho,\hatt)}{\dd \rho}\\
\begin{split}
\frac{\dd \hat{u}_z(\hatz,\rho,\hatt)}{\dd \hatz} - \frac{1}{\vareps_R} \frac{\dd R_i(z,t)}{\dd \hatz} \frac{\dd \hat{u}_z(\hatz,\rho,\hatt)}{\dd \rho} &+\\
\frac{1}{R_i(z,t) + \vareps_R \rho} & \frac{1}{\vareps_R} \frac{\dd}{\dd \rho} \left( \left( R_i(z,t) + \vareps_R \rho \right) \cdot \hat{u}_r(\hatz,\rho,\hatt) \right) = 0,
\end{split}
\end{gather}
while the boundary conditions (Equation~\ref{eq:ri} and Equation~\ref{eq:ro}) become
\begin{gather}
\frac{\dd R_i(z,t)}{\dd \hatt} = \hat{u}_{\rho}(\hatz,\rho,\hatt) - \hat{u}_{\hatz}(\hatz,\rho,\hatt) \frac{\dd R_i(z,t)}{\dd \hatz} \qquad \text{on } \rho = 0\\
\begin{split}
\frac{\dd R_i(z,t)}{\dd \hatt} + \vareps_R \frac{\dd h(z,t)}{\dd \hatt} &=\\
\hat{u}_{\rho}(\hatz,\rho,\hatt) &- \hat{u}_{\hatz}(\hatz,\rho,\hatt) \left( \frac{\dd R_i(z,t)}{\dd \hatz} + \vareps_R \frac{\dd h(z,t)}{\dd \hatz} \right) \qquad \text{on } \rho = h(z,t).
\end{split}
\end{gather}
The next step is to look for a solution in powers of the small parameter $\vareps_R \ll 1$ by writing
\[
p = p_0(\hatz,\hatt) + \vareps_R p_1(\hatz,\rho,\hatt) + \mathcal{O}(\vareps_R^2)
\]
such that the velocity components become
\begin{gather}
\hat{u}_{\hatz}(\hatz,\rho,\hatt) = -K(p_{0\hatz} - \frac{\dd R_i(z,t)}{\dd \hatz} p_{1\rho}) \left( \frac{\dd p_0(\hatz,\hatt)}{\dd \hatz} - \frac{\dd R_i(z,t)}{\dd \hatz} \frac{\dd p_1(\hatz,\rho,\hatt)}{\dd \rho} \right)\label{eq:uz_leading}\\
\hat{u}_{\rho}(\hatz,\rho,\hatt) = - \frac{1}{\vareps_L^2} K(p_{0\hatz} - \frac{1}{\vareps_R} \frac{\dd R_i(z,t)}{\dd \hatz} p_{1\rho}) \frac{\dd p_1(\hatz,\rho,\hatt)}{\dd \rho}\label{eq:ur_leading}.
\end{gather}
Applying this substitution to the governing equations and boundary conditions and letting $\vareps_R \ra 0$ leads to a simplified model in terms of the leading order pressure term $p_0(\hatz,\hatt)$. The governing equations are now Equations~\ref{eq:uz_leading}, \ref{eq:ur_leading} and the continuity equation
\beq
\frac{\dd \hat{u}_{0\hatz}(\hatz,\hatt)}{\dd \hatz} - \frac{\dd R_i(z,t)}{\dd \hatz} \frac{\dd \hat{u}_{1\hatz}(\hatz,\rho,\hatt)}{\dd \rho} + \frac{\dd \hat{u}_{1\rho}(\hatz,\rho,\hatt)}{\dd \rho} + \frac{\hat{u}_{0\rho}(\hatz,\hatt)}{R_i(z,t)} = 0\label{eq:continuity_leading}
\eeq
and the boundary conditions become
\begin{gather}
\frac{\dd R_i(z,t)}{\dd \hatt} = \hat{u}_{0\rho}(\hatz,\hatt) - \hat{u}_{0\hatz}(\hatz,\hatt) \frac{\dd R_i(z,t)}{\dd \hatz} \qquad \text{on } \rho = 0\label{eq:ri_leading}\\
\frac{\dd h(z,t)}{\dd \hatt} = \hat{u}_{1\rho}(\hatz,\rho,\hatt) - \hat{u}_{1\hatz}(\hatz,\rho,\hatt) \frac{\dd h(z,t)}{\dd \hatz} \qquad \text{on } \rho = h(z,t)\label{eq:ro_leading}.
\end{gather} 
To obtain a solution Equation~\ref{eq:continuity_leading} is integrated over the width of the BM \wrt $\rho$
\[
\int_0^{h(z,t)} \left( \frac{\dd \hat{u}_{0\hatz}(\hatz,\hatt)}{\dd \hatz} - \frac{\dd R_i(z,t)}{\dd \hatz} \frac{\dd \hat{u}_{1\hatz}(\hatz,\rho,\hatt)}{\dd \rho} + \frac{\dd \hat{u}_{1\rho}(\hatz,\rho,\hatt)}{\dd \rho} + \frac{\hat{u}_{0\rho}(\hatz,\hatt)}{R_i(z,t)} \right) d\rho = 0
\]
and yields
\beq
h(z,t) \left( \frac{\dd \hat{u}_{0\hatz}(\hatz,\hatt)}{\dd \hatz} + \frac{\hat{u}_{0\rho}(\hatz,\hatt)}{R_i(z,t)} \right) + \left[ \hat{u}_{1\rho}(\hatz,\rho,\hatt) - \frac{\dd R_i(z,t)}{\dd \hatz} \hat{u}_{1\hatz}(\hatz,\rho,\hatt) \right]_0^{h(z,t)} = 0.\label{eq:model_almost}
\eeq
The remaining terms in Equation~\ref{eq:model_almost} can be evaluated using the boundary conditions Equations~\ref{eq:ri_leading} and \ref{eq:ro_leading}. Dropping subscripts the model equation is 
\beq
\frac{\dd}{\dd \hatt} \left( R_i(z,t) \cdot h(z,t) \right) + \frac{\dd}{\dd \hatz} \left( R_i(z,t) \cdot h(z,t) \cdot u(\hatz,\hatt) \right) = 0
\eeq
with 
\[
u = -K(p_{\hatz}) \frac{\dd p(\hatz,\hatt)}{\dd \hatz}
\]

\section{Stress and strain in the artery wall}

All calculations are carried out neglecting body forces such as gravity. Under this assumption stress can only be induced in a solid by applying a force to one of its boundaries. Stress is measured in units of force per area (\SI{}{\newton\per\square\metre}, more commonly denoted as \SI{}{\pascal}). Strain, on the other hand, is a dimensionless quantity that describes the amount of deformation of a solid. Usually, one is interested in strain due to stress, \ie the amount of deformation of a solid due to some force applied to one or more of its boundaries. This relationship between stress and strain is captured in the stress-strain curve, which is unique for every material.

Here, strain is evaluated as the result of stress on the inner boundary of the artery wall. Stress is induced by the pressure pulse running through the artery and is directly proportional to pressure. First, a derivation for obtaining radial stress at any point inside the artery wall is provided. Following this derivation, a fluid-filled annulus representing the BM is introduced inside the artery wall. It is shown that for a sufficiently small BM the fluid-filled annulus can be shrunk to an infinitesimally thin ring. Due to this observation the pressure of ISF inside the BM can be directly obtained from the stress in the artery wall. This pressure serves as an input to the model of the BM derived in Supplementary Section~1.

\subsection{Plane Strain in an Annulus}
\label{sec:planestrain}

Consider an annulus (Supplementary Figure~\ref{fig:annulus}) in the $(r, \theta)$ plane of a cylindrical coordinate system on whose inner boundary ($r = a$) some pressure $P$ is applied. This corresponds to a cross-section of an idealised artery whose centre is aligned with the $z$-axis of the coordinate system. For simplicity $P$ is assumed to be continuous and uniform everywhere on the inner boundary of the annulus. Cauchy's momentum equations \cite{Howell2008}, with displacement $\disp$ independent of $z$ and $t$, are
\begin{gather}
0 = \frac{1}{r} \frac{\dd}{\dd r} \left( r \taurr \right) + \frac{1}{r} \frac{\dd \taurt}{\dd \theta} - \frac{\tautt}{r}\label{eq:cauchy1}\\
0 = \frac{1}{r} \frac{\dd}{\dd r} \left( r \taurt \right) + \frac{1}{r} \frac{\dd \tautt}{\dd \theta} - \frac{\tautt}{r}.\label{eq:cauchy2}
\end{gather}
The components of the stress tensor $\tau$ are
\begin{gather}
\taurr = (\lambda_L + 2\mu_L) \err + \lambda_L \ett \label{eq:taurr}\\
\tautt = \lambda_L \err + (\lambda_L + 2\mu_L) \err \label{eq:tautt}\\
\taurt = 2\mu_L \ert \label{eq:taurt}
\end{gather}
with material specific Lam\'e coefficients $\lambda_L$ and $\mu_L$ and strain tensor $e$ whose components are
\begin{gather}
\err = \frac{\dd v_r}{\dd r}\\
\ett = \frac{1}{r} \left( \frac{\dd v_{\theta}}{\dd \theta} + v_r \right)\\
\ert = \frac{1}{r} \frac{\dd v_r}{\dd \theta} + \frac{\dd v_{\theta}}{\dd r} - \frac{v_{\theta}}{r}.
\end{gather}
As mentioned before pressure is directly proportional to stress and therefore the boundary conditions
\begin{gather}
\taurr = -P \qquad \text{on } r = a\label{eq:boundary_p}\\
\taurr = 0 \qquad \text{on } r = b\label{eq:boundary_b}
\end{gather}
apply. Equation~\ref{eq:boundary_b} is often also referred to as a zero traction boundary condition. Two more boundary conditions are obtained from the fact that the annulus is considered to be radially symmetric and therefore
\begin{gather}
\taurt = 0 \qquad \text{on } r = a\label{eq:boundary_a}\\
\taurt = 0 \qquad \text{on } r = b.\label{eq:boundary_rtb}
\end{gather}

Equation~\ref{eq:cauchy1} and Equation~\ref{eq:cauchy2} and the boundary conditions Equation~\ref{eq:boundary_p}, Equation~\ref{eq:boundary_a} and Equation~\ref{eq:boundary_b} are nondimensionalised using the dimensionless parameters in Supplementary Table~\ref{tab:dimless}
\begin{gather}
0 = \frac{\taurr}{r} + \frac{\dd \taurr}{\dd r} + \frac{\vareps_L}{r} \frac{\dd \taurt}{\dd \theta} - \frac{\tautt}{r}\label{eq:cauchy1_nodim}\\
0 = \frac{\taurt}{r} + \frac{\dd \taurt}{\dd r} + \frac{1}{\vareps_L r} \frac{\dd \tautt}{\dd \theta} - \frac{\tautt}{r}.\label{eq:cauchy2_nodim}\\
\taurr = (1 + 2\vareps_L) \frac{\dd v_r}{\dd r} + \frac{1}{r} \frac{\dd v_{\theta}}{\dd \theta} + \frac{v_r}{r}\label{eq:taurr_nodim}\\
\tautt = \frac{\dd v_r}{\dd r} + (1 + 2\vareps_L) \left( \frac{1}{r} \frac{\dd v_{\theta}}{\dd \theta} + \frac{v_r}{r} \right)\\
\taurt = \frac{1}{r} \frac{\dd v_r}{\dd \theta} + \frac{\dd v_{\theta}}{\dd r} - \frac{v_{\theta}}{r}\label{eq:taurt_nodim}\\
\taurr = -P \qquad \text{on } r = a\\
\taurt = 0 \qquad \text{on } r = a\\
\taurr = \taurt = 0 \qquad \text{on } r = b.
\end{gather}
Expanding the displacement and stress tensor components in terms of the small parameter $\vareps_L$ leads to
\begin{gather*}
v_r = v_{r,0} + \vareps_L v_{r,1} + \mathcal{O}(\vareps_L^2)\\
v_{\theta} = v_{\theta,0} + \vareps_L v_{\theta,1} + \mathcal{O}(\vareps_L^2)\\
\taurr = \taurrl + \vareps_L \taurre + \mathcal{O}(\vareps_L^2)\\
\tautt = \tauttl + \vareps_L \tautte + \mathcal{O}(\vareps_L^2)\\
\taurt = \taurtl + \vareps_L \taurte + \mathcal{O}(\vareps_L^2).\\
\end{gather*}
It is now shown that the leading order terms $\taurrl$ and $\taurtl$ vanish. This follows from the near incompressibility of the arterial wall
\beq
\nabla \cdot \disp = \frac{\dd v_{r,0}}{\dd r} + \frac{v_{r,0}}{r} + \frac{1}{r} \frac{\dd v_{\theta,0}}{\dd \theta} = 0.\label{eq:wall_incompressible}
\eeq
The leading order terms in Equation~\ref{eq:taurr_nodim} to Equation~\ref{eq:taurt_nodim} are\\
\begin{gather}
\taurrl = \frac{\dd v_{r,0}}{\dd r} + \frac{1}{r} \frac{\dd v_{\theta,0}}{\dd \theta} + \frac{v_{r,0}}{r}\label{eq:taurrl}\\
\tauttl = \frac{\dd v_{r,0}}{\dd r} + \frac{1}{r} \frac{\dd v_{\theta,0}}{\dd \theta} + \frac{v_{r,0}}{r}\label{eq:tauttl}\\
\taurtl = \frac{1}{r} \frac{\dd v_{r,0}}{\dd \theta} + \frac{\dd v_{\theta,0}}{\dd r} - \frac{v_{\theta,0}}{r}.
\end{gather}
Both $\taurrl$ and $\taurtl$ are equal to Equation~\ref{eq:wall_incompressible} and therefore indeed vanish. The governing equations Equation~\ref{eq:cauchy1_nodim} and Equation~\ref{eq:cauchy2_nodim} become
\begin{gather}
0 = \frac{\taurre}{r} + \frac{\dd \taurre}{\dd r} + \frac{\vareps_L}{r} \frac{\dd \taurtl}{\dd \theta} + \frac{\vareps_L^2}{r} \frac{\dd \taurte}{\dd \theta} - \frac{\tautte}{r}\label{eq:cauchy1_exp}\\
0 = \frac{\taurtl}{r} + \vareps_L \frac{\taurte}{r} + \frac{\dd \taurtl}{\dd r} + \vareps_L \frac{\dd \taurte}{\dd r} + \frac{1}{\vareps_L r} \frac{\dd \tautte}{\dd \theta} - \frac{\taurtl}{r} - \vareps_L \frac{\taurte}{r}\label{eq:cauchy2_exp}
\end{gather}
and the boundary conditions are
\begin{gather}
\taurre = -P \qquad \text{on } r = a\\
\taurte = 0 \qquad \text{on } r = a\\
\taurre = \taurte = 0 \qquad \text{on } r = b.
\end{gather}
Letting $\vareps_L \ra 0$ and dropping subscripts Equation~\ref{eq:cauchy2_exp} vanishes completely and Equation~\ref{eq:cauchy1_exp} reduces to
\beq
0 = \frac{\taurr}{r} + \frac{d \taurr}{d r} - \frac{\tautt}{r}\label{eq:planestrain}
\eeq
with boundary conditions
\begin{gather}
\taurr = -P \qquad \text{on } r = a\label{eq:plane_bca}\\
\taurr = 0 \qquad \text{on } r = b.\label{eq:plane_bcb}
\end{gather}
The system of Equation~\ref{eq:planestrain} to Equation~\ref{eq:plane_bcb} now only depends on $r$. This makes sense as pressure is applied uniformly across the boundary on $r = a$ and therefore does not vary in the $\theta$-direction. To solve for $\taurr$ an Airy stress function $\Airy(r)$ is used that satisfies the biharmonic equation
\beq
\nabla^4 \Airy(r) = 0.\label{eq:biharmonic}
\eeq
Similar to a stream function in fluid dynamics the Airy stress function is a potential function for the stresses in a solid and is only defined to within a linear function in $r$ \cite{Howell2008}. Again, for readability the notation $\Airy = \mathcal{A}(r)$ is used from now on. To find a solution for $\Airy$ one can begin by defining
\beq
\taurr = \frac{1}{r} \frac{d \Airy}{dr}
\eeq
and from Equation~\ref{eq:planestrain}
\beq
\tautt = \frac{d^2 \Airy}{dr^2}
\eeq
follows. Expanding Equation~\ref{eq:biharmonic} the biharmonic equation for $\Airy$ is
\beq
\frac{1}{r} \frac{d}{dr} \left( r \frac{d}{dr} \left( \frac{1}{r} \frac{d}{dr} \left( r \frac{d\Airy}{dr} \right) \right) \right) = 0,
\eeq
which has a solution of the form
\beq
\Airy = k_1 r^2 + k_2 + k_3 r^2 \log(r) + k_4 \log(r).\label{eq:airy}
\eeq
To solve for the unknown constants $k_1, \ldots, k_4$ the boundary conditions Equation~\ref{eq:plane_bca} and Equation~\ref{eq:plane_bcb} are used. At $r = a$ the stress on the boundary is determined by pressure $P$ such that
\beq
\frac{1}{r} \frac{d\Airy}{dr} = -P \qquad \text{on } r = a.\label{eq:airy_bca}
\eeq 
Equation~\ref{eq:plane_bcb} means that no surface traction is applied on the boundary $r = b$. Surface traction $t(s)$ of arc length $s$ is given by
\beq
t(s) = \frac{d}{ds} \left( \frac{d\Airy}{dr} \right).
\eeq
Integrating \wrt $s$ and applying $t(s) = 0$ gives
\beq
\frac{d\Airy}{dr} = C \qquad \text{on } r = b\label{eq:airy_bcb_c},
\eeq
where $C$ is the constant of integration. Remembering that $\Airy$ is only defined to within a linear function in $r$ one can, without loss of generality, define $C = 0$. Thus, Equation~\ref{eq:airy_bcb_c} becomes
\beq
\frac{d\Airy}{dr} = 0 \qquad \text{on } r = b\label{eq:airy_bcb}.
\eeq
To obtain a condition for $\Airy$ on the boundary $r = b$ rewrite
\beq
\frac{d\Airy}{dr} = \frac{d\Airy}{ds} \vect{e}_t + \frac{d\Airy}{dn} \vect{e}_n,
\eeq
where $\vect{e}_t$ and $\vect{e}_n$ are the unit tangent and outer normal vectors on the boundary. Integrating \wrt $s$ and using Equation~\ref{eq:airy_bcb} yields
\beq
\Airy = 0 \qquad \text{on } r = b.\label{eq:airy_bcb2}
\eeq
Thus far there are three boundary conditions (Equation~\ref{eq:airy_bca}, Equation~\ref{eq:airy_bcb} and Equation~\ref{eq:airy_bcb2}) for the four unknown constants in Equation~\ref{eq:airy}. The fourth condition required to solve Equation~\ref{eq:airy} is called a compatibility condition and is necessary to ensure that the displacement is unique for a certain strain. This condition is not automatically satisfied since there are two stress components $\taurr$ and $\tautt$ for just one displacement component $v_r(r)$. Differentiating $\taurr$ and $\tautt$ and adding them yields the compatibility condition
\beq
\frac{d^3 \Airy}{dr^3} + \frac{1}{r} \frac{d^2 \Airy}{dr^2} - \frac{1}{r^2} \frac{d\Airy}{dr} = 0.
\eeq
which is satisfied only if $k_3 = 0$. Using the boundary conditions to determine the remaining constants results in the solution
\beq
\Airy(r) = \frac{P a^2 b^2}{a^2 - b^2} \cdot \log \left( \frac{r}{b} \right) + \frac{P a^2}{2 (a^2 - b^2)} \left( b^2 - r^2 \right)
\eeq
for a given radial position $r$. The stress components evaluate to
\begin{gather}
\taurr = \frac{P a^2}{b^2 - a^2} \left( 1 - \frac{b^2}{r^2} \right)\label{eq:taurr_airy}\\
\tautt = \frac{P a^2}{b^2 - a^2} \left( 1 + \frac{b^2}{r^2} \right).
\end{gather}
The Airy stress function approach provides a solution for a material with a linear stress-strain relation.

\subsection{Validation of the Plane Strain Model}

The plane strain model for calculating the stresses inside an artery makes a number of simplifying assumptions:
\begin{enumerate}
\item Stress is calculated without regard for the longitudinal $z$-direction.
\item The artery wall is a linear elastic material.
\item Pressure inside the artery is uniform across the lumen area.
\end{enumerate}
The validity of these assumptions shall be addressed in this section before proceeding with an adaptation of the plane strain model to obtain ISF pressure inside the BM.

The first assumption arises from the fact that blood vessels are longitudinally tethered, \ie both ends of the arterial tree are attached to the heart, which limits the amount of stretch they can undergo along the $z$-direction. This assumption has been used in the literature and the same equations have successfully been used to accurately describe blood flow dynamics by others \cite{Olufsen2000,Kolachalama2007,Devault2008,Cousins2014}. To evaluate the second assumption one has to recognise that the artery wall actually is a viscoelastic material. This means that there is a time delay between an increase in blood pressure (\ie increased radial stress on the artery wall) and the corresponding increase in vessel diameter. This effect cannot accurately be captured by a purely elastic material. However, \cite{Tardy1991} showed that this time lag is mainly an experimental artefact and have developed a model to correct experimental measurements for this artefact. No other publication could be found to investigate this hypothesis, but results from other blood flow simulation studies show that this simple model appears to suffice to accurately model blood flow dynamics.

To confirm the second and address the third assumption a simple model was set up in the finite element multiphysics modelling software COMSOL 5.1. The geometry is a 2D axisymmetric tube of length $L_0$ with radius $R_0$ surrounded by a wall of thickness $h_0$. Flow inside the tube is governed by the Navier-Stokes equations for laminar flow
\begin{gather}
\rho (\vel \cdot \nabla) \vel = \nabla \cdot \left( -p \boldsymbol{I} + \mu \left( \nabla \vel + (\nabla \vel)^T \right) \right) + \boldsymbol{F}_s\\
\rho \nabla \cdot \vel = 0
\end{gather}
with $\boldsymbol{I}$ identity matrix and $\boldsymbol{F}_s$ some external body force. The dynamics of the wall are governed by either the equations for linear elasticity or hyperelasticity. In both cases the stress relates to $\boldsymbol{F}_s$ via
\beq
-\nabla \sigma = \boldsymbol{F}_s \nu_s\\
\eeq
with $\sigma$ stress tensor and $\nu_s$ Poisson's ratio. In the case of the linear elastic model stress is governed by Hooke's law
\begin{gather}
\sigma = \sum_i \sum_j C_{ij} \vareps_{\text{el},ji}\\
\vareps_{\text{el}} = \frac{1}{2} \left( (\nabla \disp)^T + \nabla \disp + (\nabla \disp)^T \nabla \disp \right) 
\end{gather}
with $C$ elasticity tensor and $\vareps_{\text{el}}$ elastic strain tensor. In the case of nonlinear elasticity stress is governed by a Neo-Hookean material model, which has been shown to provide a good fit for the mechanical response of arterial elastin \cite{Holzapfel1998,Gundiah2007}, where stress is given as
\beq
\sigma = J^{-1} \boldsymbol{F}_s \boldsymbol{S}_s \boldsymbol{F}_s^T
\eeq
with $J = \det(\boldsymbol{F}_s)$ and $\boldsymbol{S}_s = \dd W_s / \dd \vareps_{el}$ second Piola-Kirchoff stress. For a nearly incompressible neo-Hookean material the strain energy density function is
\beq
W_s = W_{\text{iso}} + W_{\text{vol}} = \frac{1}{2} \mu_s (\bar{I}_1 - 3) + \frac{1}{2} K_s (J_{\text{el}} - 1)^2,
\eeq
where $\mu_s$ initial shear modulus, $\bar{I}_1$ first principal invariant of the symmetric right Cauchy-Green tensor, $K_s$ initial bulk modulus and $J_{\text{el}}$. The boundary between the lumen and artery wall implements a two-way coupling between the two domains: In the fluid domain pressure is determined by the normal stress on the wall
\beq
\sigma \cdot \norm = \left( -p \boldsymbol{I} + \mu \left( \nabla \vel + (\nabla \vel)^T \right) \right) \cdot \norm,
\eeq
while the rate of displacement of the artery wall is determined by the fluid velocity
\beq
\frac{\dd \disp}{\dd t} = \vel.
\eeq
Because the artery is tethered displacement of the artery wall is restricted to the $z$-direction at open ends
\beq
\disp \cdot \norm = 0.
\eeq
At the fluid inlet a velocity is applied, while at the outlet pressure is applied. The parameter values used for the simulations are listed in Supplementary Table~\ref{tab:airy_sim}.

The results of this simulation are shown in Supplementary Figure~\ref{fig:linear_hyperelastic}. There is good agreement between the hyperelastic and linear elastic material while the Airy stress function overestimates hoop stress ($\tau_{\theta\theta}$) at the higher end of the pressure range. This is acceptable as an overestimation of stress would lead to an overestimation of ISF pressure inside the BM. If ISF pressure inside the BM is too small to drive significant flow when stress is estimated using the Airy stress function then it will also be too small when using a hyperelastic numerical simulation. Furthermore, only radial stress is required for the calculation of ISF pressure and the overall agreement between the Airy stress function and the simulation results is found to be suitably good for radial stress. Therefore the Airy stress function calculation was used for the remainder of this thesis.

\subsection{Plane Strain in an Artery Cross-Section with BM}
\label{sec:bm_stress}

The Airy stress function derived in Section~\ref{sec:planestrain} can also be used to determine the pressure of ISF in the BM. For simplicity the BM shall be treated as a thin fluid-filled empty space. This approximation is considered valid for the fluid part of a porous medium if the fluid-filled space is small enough. The BM is only approximately \SI{150}{\nano\metre} in thickness, which is roughly three orders of magnitude smaller than the thickness of the artery wall of the MCA and therefore, the approximation is considered valid. Two independent Airy stress functions $\Airy_1(r)$ and $\Airy_2(r)$ are thus used to describe the artery wall parts surrounding the BM (Supplementary Figure~\ref{fig:twoannuli}). Following the analysis in Section~\ref{sec:planestrain} these are
\begin{gather}
\Airy_1 = k_1 r^2 + k_2 + k_3 \log(r)\\
\Airy_2 = k_4 r^2 + k_5 + k_6 \log(r)
\end{gather}
and as before the boundary conditions are
\begin{gather}
\frac{1}{r} \frac{d \Airy_1}{dr} = -P \qquad \text{on } r = a_1\\
\Airy_2 = \frac{d \Airy_2}{dr} = 0 \qquad \text{on } r = b_2.
\end{gather}
The pressure of ISF $P_{BM}$ is required to be continuous and uniform and therefore
\beq
\frac{d \Airy_1}{dr} = \frac{d \Airy_2}{dr} \qquad \text{on } r = b_1, a_2
\eeq
applies.

The claim that, given a sufficiently thin BM, a single Airy stress function can be used to model the artery wall like in Section~\ref{sec:planestrain} and the ISF pressure can then be directly obtained from the stress in the wall at the location of the BM shall be proven here. Because ISF is an incompressible fluid its area $A_f$ has to remain constant
\beq
A_f = \pi \left( \left( a_2 + v_r(a_2) \right)^2 - \left( b_1 + v_r(b_1) \right)^2 \right) = const.
\eeq
Eliminating $\pi$ and the small square terms in $v_r$ yields
\beq
2a_2 v_r(a_2) - 2b_1 v_r(b_1) = b_1^2 - a_2^2.
\eeq
As the fluid part of the BM is considered to be very thin one can take the limit
\beq
\lim_{b_1 \ra a_2} 2a_2 v_r^+(a_2) - 2b_1 v_r^-(b_1) = b_1^2 - a_2^2.
\eeq
and confirm
\beq
v_r^+(a_2) = v_r^-(b_1).
\eeq
Therefore $\Airy_1$ and $\Airy_2$ must be continuous across the BM such that
\beq
\Airy_1 = \Airy_2 \qquad \text{on } r = b_1, a_2.
\eeq
Therefore the claim is proven and a single Airy stress function $\Airy = \Airy_1 + \Airy_2$ can be used to describe the artery wall. The ISF pressure can then be found by evaluating $\taurr$ (Equation~\ref{eq:taurr_airy}) at $r = b_1$
\beq\label{eq:bm_pressure}
P_{\text{BM}} = - \frac{P a_1^2}{b_2^2 - a_1^2} \left( 1 - \frac{b_2^2}{b_1^2} \right).
\eeq

\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
\begin{thebibliography}{10}
\bibitem{Howell2008}
Howell~P, Kozyreff~G and Ockendon~J (2008) {\em Applied Solid Mechanics.} Cambridge University Press, Cambridge, UK.

\bibitem{Olufsen2000}
Olufsen~MS et~al. (2000) Numerical Simulation and Experimental Validation of Blood Flow in Arteries with Structured-Tree Outflow Conditions. {\em Ann.~Biomed.~Eng.} 28(11):1281--1299.

\bibitem{Kolachalama2007}
Kolachalama~V, Bressloff~NW, Nair~PB, Shearman~CP (2007) Predictive Haemodynamics in a One-Dimensional Carotid Artery Bifurcation. Part I: Application to Stent Design. {\em IEEE~Trans.~Biomed.~Eng.} 54(5):802--812.

\bibitem{Cousins2014}
Cousins~W, Gremaud~PA (2014) Impedance boundary conditions for general transient hemodynamics. {\em Int.~J.~Numer.~Method~Biomed.~Eng.}

\bibitem{Devault2008}
Devault~K and Gremaud~A (2008) Blood Flow in the Circle of Willis: Modeling and Calibration. {\em Multiscale~Model.~Simul.} 7(2):888--909.

\bibitem{Tardy1991}
Tardy~Y, Meister~JJ, Perret~F, Brunner~HR and Arditi~M (1991) Non-invasive estimate of the mechanical properties of peripheral arteries from ultrasonic and photoplethysmographic measurements. {\em Clin.~Phys.~Physiol.~Meas.} 12(1):39--54.

\bibitem{Holzapfel1998}
Holzapfel~GA and Weizsacker~HW (1998) Biomechanical behaviour of the arterial wall and its numerical characterization. {\em Comput.~Biol.~Med.} 28(4):377--392.

\bibitem{Gundiah2007}
Gundiah~N, Ratcliffe~MB and Pruitt~LA (2007) Determination of strain energy function for arterial elastin: Experiments using histology and mechanical tests. {\em J.~Biomech.} 40(3):586--594.  
\end{thebibliography}

\begin{table}[h]
\textbf{\refstepcounter{table}\label{tab:dimless} Supplementary Table \arabic{table}.}{ Dimensionless variables and parameters used to nondimensionalise the equations of the stress tensor components of an inflated annulus (Equation~\ref{eq:taurr}, Equation~\ref{eq:tautt} and Equation~\ref{eq:taurt}). }

\processtable{ }
{\begin{tabular}{cc}\toprule
Parameter & Physical meaning \\\midrule
$L$         & length\\
$R$         & radius\\
$V$         & tangential displacement\\
$\lambda_L$ & first Lam\'e coefficient\\
$\mu_L$ & second Lam\'e coefficient\\
$\vareps_L = R/L \ll 1$ & radius-to-length ratio\\
\midrule
Variable       & \\
\midrule
$r \sim R = \vareps_L L$ & radius\\
$v_{r} \sim \vareps_L V$ & radial displacement\\
$v_{\theta} \sim V$ & tangential displacement\\
$\taurr \sim \lambda_L V/L$ & radial stress\\
$\tautt \sim \lambda_L V/L$ & hoop stress\\
$\taurt \sim \mu_L V/L$ & \\
$P \sim \lambda_L V/L$ & blood pressure\\\botrule
\end{tabular}}{}
\end{table}

\begin{table}[h]
\textbf{\refstepcounter{table}\label{tab:airy_sim} Supplementary Table \arabic{table}.}{ Parameters used in the linear and hyperelastic FE simulation of an artery. }

\processtable{ }
{\begin{tabular}{ccc}\toprule
Parameter & Value & Physical meaning \\\midrule
$R$ & \SI{0.134}{\centi\metre} & vessel radius\\
$L$ & \SI{1.2}{\centi\metre} & vessel length\\
$h$ & $R/4$ & wall thickness\\
$u_\text{in}$ & \SI{0.6}{\centi\metre\per\second} & inlet velocity\\
$p_\text{out}$ & 80 / 100 / \SI{120}{\mmHg} & outlet pressure\\
$\mu$ & \SI{4.88e-3}{\pascal\second} & dynamic viscosity of blood\\
$\rho$ & \SI{1.06}{\gram\per\cubic\metre} & blood density\\
$\rho_s$ & \SI{0.9}{\gram\per\cubic\metre} & artery wall density\\
\midrule
\multicolumn{3}{l}{Linear elastic material properties}\\
\midrule
$E$ & \SI{7.37e5}{Pa} & Young's modulus\\
$\nu_s$ & 0.495 & Poisson's ratio\\
\midrule
\multicolumn{3}{l}{Hyperelastic material properties}\\
\midrule
$\mu_s$ & $E / (2 (1 + \nu))$ & initial shear modulus\\
$K_s$ & $E / (3 (1 - 2\nu))$ & initial bulk modulus\\\botrule
\end{tabular}}{}
\end{table}

\begin{figure}
\begin{center}
\includegraphics[width=15cm]{figures/annulus.pdf}% This is a *.jpg file
\end{center}
\textbf{\refstepcounter{figure}\label{fig:annulus} Supplementary Figure \arabic{figure}.}{ Schematic representation of a linear elastic annulus in the $(r, \theta)$ plane inflated by an internal pressure $P$. }
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=15cm]{figures/linear_hyperelastic.png}% This is a *.jpg file
\end{center}
\textbf{\refstepcounter{figure}\label{fig:linear_hyperelastic} Supplementary Figure \arabic{figure}.}{ Comparison of radial and hoop stresses in an artery wall using the Airy stress function (black line) or an FE simulation with linear (green circles) or hyperelastic (blue squares) material at various blood pressures. (a) $p = \SI{80}{\mmHg}$, (b) $p = \SI{100}{\mmHg}$, (c) $p = \SI{120}{\mmHg}$. }
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=15cm]{figures/twoannuli.pdf}% This is a *.jpg file
\end{center}
\textbf{\refstepcounter{figure}\label{fig:twoannuli} Supplementary Figure \arabic{figure}.}{ Schematic representation of the elastic model of an artery cross-section with BM inflated by an internal pressure $P$. For simplicity the BM is approximated by a fluid-filled space at some radial position $r$. The artery wall is then split into two separate parts, which are described by two independent Airy stress functions $\Airy_1$ and $\Airy_2$. }
\end{figure}

\end{document}

### 2017-07-29 paper.tex
reviewers comments

@@
 
-\title[Arterial pulsations cannot drive periarterial drainage]{Arterial pulsations cannot drive intramural periarterial drainage: Significance for Alzheimer's disease}
+\title[Arterial pulsations cannot drive periarterial drainage]{Arterial pulsations cannot drive intramural periarterial drainage: Significance for \Ab drainage}
 
@@
 No significant differences were found between mice that received the pulse modulator and control mice.
+We were unable to measure whether the treatment of mice did not, in addition to the desired reduction in pulsation frequency, lead to a decrease in the magnitude of pulsations.
+However, given that the flow rate calculated from the computational model is eight orders of magnitude smaller than what would have been expected it is highly unlikely that the administration of the beta-blocker caused a pressure drop large enough to solely cause this discrepancy in flow rates.
 It appears that other forces are neccessary to produce the pressure gradients required to push fluid through the BM and an interesting candidate for such forces is the contraction of vascular smooth muscle cells, as noted in \cite{DiMarco2015} and modelled in detail by \cite{Aldea2016}.

### 2017-07-29 paper.tex
modified introduction for reviewer 2

@@
 The protein amyloid-\textbeta\xspace (\Ab) is produced as part of the normal metabolism of healthy brains, but it plays a major role in the development of AD \cite{Haass1992}.
-Its accumulation in the walls of cerebral blood vessels as cerebral amyloid angiopathy (CAA) is closely associated with the onset and progression of AD.
-CAA strongly correlates with the progression of AD and the degree of dementia as it appears in \SIrange{90}{96}{\percent} of AD patients \cite{Weller2006, Weller2009}.
+According to the \Ab hypothesis AD is caused by a failure to remove \Ab from the brain and successively, its accumulation in the form of plaques in the parenchyma.
+Prior to the buildup of plaques in the brain \Ab accumulates in the walls of cerebral blood vessels as cerebral amyloid angiopathy (CAA), which correlates much better with the degree of dementia compared to the number and size of plaques and appears in \SIrange{90}{96}{\percent} of AD patients \cite{Weller2006, Weller2009}.
 
@@
 While the rest of the body possesses a dedicated network of lymphatic vessels these do not exist in the brain \cite{Weller2010}.
-There is ample evidence that the basement membranes (BM) in the walls of cerebral arteries provide the clearance pathway for interstitial fluid (ISF) towards the surface of the brain, also referred to as intramural periarterial drainage (IPAD) \cite{Weller2010,Carare2008,Hawkes2011,Morris2014}.
+Both the pathology of CAA and ample evidence from injection studies suggest that the basement membranes (BM) in the walls of cerebral arteries provide the clearance pathway for interstitial fluid (ISF) towards the surface of the brain, also referred to as intramural periarterial drainage (IPAD) \cite{Weller2010,Carare2008,Hawkes2011,Morris2014}.
 It is important not to confuse the IPAD pathway with the Virchow-Robin spaces, which are often referred to as perivascular (or ``paravascular'') spaces, possibly artefactual, located between the glia limitans and the arterial wall.

### 2017-07-29 paper.tex
added injection location

@@
 
-\subsection{Perivascular Drainage through the MCA}
+\subsection{IPAD through the MCA}
 
@@
 Baseline measurements of oxygen saturation, heart rate, breath rate and pulse distension were taken before administering Ate.
-A volume of \SI{0.5}{\micro\litre} of \Ab$_{1-40}$ HiLyte Fluor 555 (Cambridge Bioscience, UK) was stereotaxically injected into the left hippocampus at a rate of \SI{0.2}{\micro\litre\per\minute}.
+A volume of \SI{0.5}{\micro\litre} of \Ab$_{1-40}$ HiLyte Fluor 555 (Cambridge Bioscience, UK) was stereotaxically injected into the left hippocampus at a rate of \SI{0.2}{\micro\litre\per\minute} (coordinates from Bregma: AP = \SI{-1.9}{\mm}; ML = \SI{1.5}{\mm} and DV = \SI{-1.7}{\mm}).
 Injection pipettes were left \textit{in situ} for \SI{2}{\minute} to prevent reflux.

### 2017-08-03 paper.tex
added Maureen as author
changed order of authors
modifications for final version

@@
 \def\firstAuthorLast{Diem {et~al.}} %use et al only if is more than 1 author
-\def\Authors{Alexandra K. Diem\,$^{1,*}$, Neil W. Bressloff\,$^{1}$, Roxana O. Carare\,$^{2}$, Matthew MacGregor Sharp\,$^{2}$ and Giles Richardson\,$^3$}
+\def\Authors{Alexandra K. Diem\,$^{1,*}$, Matthew MacGregor Sharp\,$^{2}$, Maureen Gatherer\,$^{2}$, Neil W. Bressloff\,$^{1}$, Roxana O. Carare\,$^{2}$ and Giles Richardson\,$^3$}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
 
-In order to process metabolites, the brain requires a lymphatic system that transports waste products to the lymph nodes, where they are decomposed.
+In order to process metabolites, the brain requires a lymphatic system that transports waste products to the lymph nodes.
 While the rest of the body possesses a dedicated network of lymphatic vessels these do not exist in the brain \cite{Weller2010}.
@@
 It is important not to confuse the IPAD pathway with the Virchow-Robin spaces, which are often referred to as perivascular (or ``paravascular'') spaces, possibly artefactual, located between the glia limitans and the arterial wall.
-Virchow-Robin spaces are virtual spaces and have no role in \Ab drainage.
+Virchow-Robin spaces are compartments occupied by the pial-glial basement membranes spaces and have no role in \Ab drainage from the extracellular spaces.
 The BM forming the IPAD does not actually constitute a space, but rather a fluid-filled protein matrix.
@@
 Brains were cut into \SI{20}{\micro\metre} coronal sections using a freezing microtome (Leica) and the injection site located by examining on a Zeiss Axioskop 2 fitted with a rhodamine filter.
-Sections of \SI{400}{\micro\metre} and \SI{800}{\micro\metre} thickness both posterior and anterior to the injection site were selected for immunohistochemistry.
+Sections \SI{400}{\micro\metre} and \SI{800}{\micro\metre} both posterior and anterior to the injection site were selected for immunohistochemistry.
 Sections were blocked in \SI{15}{\percent} goat serum (Sigma 9023) for \SI{1}{\hour} at room temperature (RT).
@@
 designed and analysed numerical models, R.O.C and M.M.S.
-designed and performed mouse experiments, A.K.D., N.W.B, M.M.S., R.O.C.
+designed and performed mouse experiments, M.G.
+analysed brain sections and performed image analysis, A.K.D., N.W.B, M.M.S., R.O.C.
 and G.R.
@@
 \centering
-\includegraphics[width=10cm]{figures/pvs_model.pdf}
+\includegraphics[width=10cm]{figures/fig01.jpg}
 \textbf{\refstepcounter{figure}\label{fig:pvs_model} Figure \arabic{figure}.}{Geometry of the perivascular drainage pathways model.
@@
 \centering
-\includegraphics[width=10cm]{figures/inlet.png}
+\includegraphics[width=10cm]{figures/fig02.jpg}
 \textbf{\refstepcounter{figure}\label{fig:inlet} Figure \arabic{figure}.}{Inlet function for MCA simulations obtained from patient-specific measurements used to represent one period of \SI{0.85}{\second} of the pulse wave.
@@
 \centering
-\includegraphics[width=10cm]{figures/mca.pdf}
+\includegraphics[width=10cm]{figures/fig03.jpg}
 \textbf{\refstepcounter{figure}\label{fig:mca} Figure \arabic{figure}.}{Blood pressure (a), wall displacement (b) and the resulting ISF pressure (c) in the MCA.
@@
 \centering
-\includegraphics[width=10cm]{figures/h_dh.png}
+\includegraphics[width=10cm]{figures/fig04.jpg}
 \textbf{\refstepcounter{figure}\label{fig:h} Figure \arabic{figure}.}{BM width $h_\textrm{bm}(z,t)$ at several time points during the cardiac cycle (a) and its gradient along the $z$-axis at the same time points (b).
@@
 \centering
-\includegraphics[width=10cm]{figures/valve_test.png}
+\includegraphics[width=10cm]{figures/fig05.jpg}
 \textbf{\refstepcounter{figure}\label{fig:valve_test} Figure \arabic{figure}.}{ISF flux (positive values in the direction of arterial flow) averaged over one cycle as a function of BM position $\eta$ in the arterial wall.
@@
 \centering
-\includegraphics[width=10cm]{figures/confocal.jpg}
+\includegraphics[width=10cm]{figures/fig06.jpg}
 \textbf{\refstepcounter{figure}\label{fig:confocal} Figure \arabic{figure}.}{Composite tile scan of confocal images of the distribution of intracerebrally injected \Ab, in relation to collagen IV and SMA.
+The image was taken at \SI{400}{\micro\m} anterior to the injection site in an atenolol (Ate) injected mouse, to avoid any artefacts induced by the injection tract, relying instead on the pattern of distribution of the fluorescently injected Abeta away from the direct injection.
 Immunostaining \Ab: red, SMA: green, collagen IV: blue.
-Arrows are pointing to arteries with \Ab in the intramural perivascular drainage pathways.}
+Arrows are pointing to arteries with \Ab in the intramural perivascular drainage pathways.
+Scale bars = \SI{400}{\micro\m}}
 \end{figure}
@@
 \centering
-\includegraphics[width=10cm]{figures/graph.png}
+\includegraphics[width=10cm]{figures/fig07.jpg}
 \textbf{\refstepcounter{figure}\label{fig:graph} Figure \arabic{figure}.}{Graphs showing the number of arteries, veins and capillaries with \Ab in their walls in atenolol treated mice compared to control mice.

### 2017-08-03 figures/pvs_model.py

@@ -1,55 +1,0 @@
-from __future__ import division
-from matplotlib.pyplot import plot, annotate, arrow, axhline, xlim, ylim, xlabel, ylabel, xticks, yticks, savefig, show, figure, text
-from matplotlib import rc
-import numpy as np
-import os
-
-rc('font', family='sans-serif', size=9)
-rc('text', usetex=True)
-
-pi = np.pi
-end = 80*pi
-mid = end/2
-eta = np.linspace(0, end, 40*256, endpoint=True)
-Hi = 0.05*np.sin(0.05*eta)+0.99
-Ho = 0.05*np.sin(0.05*(eta+20))+1.3
-Hb = 0.05*np.sin(0.05*(eta+5))+0.5
-
-arrow_start = mid-20
-
-def cm2inch(value):
-    return value/2.54
-
-fig = figure(figsize=(cm2inch(17.2), cm2inch(8.7)))
-
-plot(eta, Hi, linewidth=2, color='k')
-plot(eta, Ho, linewidth=2, color='k')
-plot(eta, Hb, linewidth=2, color='k')
-
-axhline(y=0.99, linewidth=1, linestyle='dashed', color='k')
-axhline(y=1.3, linewidth=1, linestyle='dashed', color='k')
-axhline(y=0.5, linewidth=1, linestyle='dashed', color='k')
-
-annotate('artery wall', (mid,0.7), xycoords='data', xytext=(-22, 0), textcoords='offset points')
-annotate('ISF flow', (mid,1.15), xycoords='data', xytext=(-17,0.05), textcoords='offset points')
-annotate('artery wall', (mid,1.4), xycoords='data', xytext=(-22,0), textcoords='offset points')
-annotate('blood flow', (mid,0.2), xycoords='data', xytext=(-20,0), textcoords='offset points')
-annotate(r'$h_\textrm{bm}(z,t)$', (50,1.11), xycoords='data', xytext=(5,0), textcoords='offset points')
-
-arrow(arrow_start, 0.15, 40, 0, head_width=0.05, head_length=3, color='k')
-arrow(arrow_start+40, 1.1, -40, 0, head_width=0.05, head_length=3, color='k')
-arrow(60, 1.09, 0, -0.06, head_width=2, head_length=0.02, color='k')
[…]
